# Detection Accuracy{#preds_detect}

To this point we have:

1. Provided a data overview: [here](#data_desc) and [here](#data_load)
2. [Processed the UAS point cloud](#ptcld_process)
3. [Demonstrated our geometry-based slash pile detection methodology](#geom_detect)
4. [Demonstrated our spectral refinement (i.e. data fusion) methodology](#data_fusion)
5. [Reviewed how we will evaluate our method](#meth_eval)
5. and [Made predictions using our method on four experimental sites](#meth_preds)

In this section, we'll evaluate the effectiveness of the proposed geometric, rules-based slash pile detection methodology by assessing its detection accuracy performance. We fully reviewed the detection accuracy assessment [workflow here](#detect_metrics_form), but here is a quick overview:

Detection accuracy metrics are calculated by aggregating raw TP (true positive), FP (false positive; commission), and FN (false negative; omission) counts to quantify the method's ability to find the piles. Aggregation of the instance matching allows us to evaluate omission rate (false negative rate or miss rate), commission rate (false positive rate), precision, recall (detection rate), and the F-score metric. As a reminder, true positive (TP) instances correctly match ground truth instances with a prediction, commission predictions do not match a ground truth instance (false positive; FP), and omissions are ground truth instances for which no predictions match (FN)

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Instance Matching

The first step in evaluating the performance of our slash pile detection methodology is to perform instance matching. Instance matching is the process of checking our predictions of slash pile presence (or absence) against the actual pile locations in real life (we use image-annotated pile outlines as ground truth) to determine if the method correctly identifies presence and absence. We use the framework established by Puliti et al. [(2023, p. 14)](https://doi.org/10.48550/arXiv.2309.01279) to validate detections based on an Intersection over Union (IoU) threshold. We set this threshold at 45% which is slightly more permissive than the 50% used by Puliti et al. (2023) for tree crowns to accommodate the fact that our target objects are located on the forest floor rather than in the canopy and are therefore subject to higher levels of occlusion from aerial data.

we reviewed this process and defined the function `ground_truth_prediction_match()` [here](#iou_match)

```{r, results=F}
# map over ground_truth_prediction_match for dbscan
  dbscan_gt_pred_match <- 
    all_stand_boundary$site_data_lab %>% 
    purrr::set_names() %>% 
    purrr::map(
      \(x)
      ground_truth_prediction_match(
          ground_truth = slash_piles_polys[[x]] %>% 
            dplyr::filter(is_in_stand) %>% 
            sf::st_transform( sf::st_crs(dbscan_spectral_preds[[x]]) ) %>% 
            dplyr::arrange(desc(image_gt_area_m2)) # this is so the algorithm starts with the largest
          , gt_id = "pile_id"
          , predictions = dbscan_spectral_preds[[x]]
          , pred_id = "pred_id"
          , min_iou_pct = 0.45
        )
      , .progress = T
    )
# map over ground_truth_prediction_match for watershed
  watershed_gt_pred_match <- 
    all_stand_boundary$site_data_lab %>% 
    purrr::set_names() %>% 
    purrr::map(
      \(x)
      ground_truth_prediction_match(
          ground_truth = slash_piles_polys[[x]] %>% 
            dplyr::filter(is_in_stand) %>% 
            sf::st_transform( sf::st_crs(watershed_spectral_preds[[x]]) ) %>% 
            dplyr::arrange(desc(image_gt_area_m2)) # this is so the algorithm starts with the largest
          , gt_id = "pile_id"
          , predictions = watershed_spectral_preds[[x]]
          , pred_id = "pred_id"
          , min_iou_pct = 0.45
        )
      , .progress = T
    )
```

let's see what the instance matching data looks like

```{r}
# what did we get?
dbscan_gt_pred_match[[1]] %>% 
  dplyr::glimpse()
```

let's quickly look at the distribution of IoU for the TP matches

```{r, fig.height=9.7}
# palette
pal_match_grp = c(
  "omission"=viridis::cividis(3)[1]
  , "commission"= viridis::cividis(3)[2]
  , "true positive"=viridis::cividis(3)[3]
)
# plot
iou_dta_temp <- 
  dplyr::bind_rows(
   dbscan_gt_pred_match %>% 
      purrr::imap(
        \(x,nm)
        x %>% 
        dplyr::filter(match_grp=="true positive") %>% 
        dplyr::mutate(
          site = all_stand_boundary %>% dplyr::filter(site_data_lab==nm) %>% dplyr::pull(site)
          , method = "dbscan"
        )
      ) %>% 
     dplyr::bind_rows()
   , watershed_gt_pred_match %>% 
      purrr::imap(
        \(x,nm)
        x %>% 
        dplyr::filter(match_grp=="true positive") %>% 
        dplyr::mutate(
          site = all_stand_boundary %>% dplyr::filter(site_data_lab==nm) %>% dplyr::pull(site)
          , method = "watershed"
        )
      ) %>% 
     dplyr::bind_rows()
  )
# plot it

iou_dta_temp %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(mapping = ggplot2::aes(x = iou)) +
  ggplot2::geom_vline(xintercept = 0.45, color = "gray", linetype = "dashed", lwd = 1.5) +
  ggplot2::geom_violin(
    mapping = ggplot2::aes(y = 0, color = match_grp, fill = match_grp)
    , alpha = 0.8
  ) +
  ggplot2::geom_boxplot(width = 0.1, fill = NA, outliers = F) +
  ggplot2::scale_color_manual(values=pal_match_grp) +
  ggplot2::scale_fill_manual(values=pal_match_grp) +
  ggplot2::scale_y_continuous(NULL,breaks=NULL) +
  ggplot2::scale_x_continuous(limits = c(0,1), labels=scales::percent, breaks = scales::breaks_extended(n=6)) +
  ggplot2::facet_grid(
    rows = dplyr::vars(site)
    , cols = dplyr::vars(method)
    # , scales = "free_y"
    , switch = "y"
    , axes = "all_x"
  ) +
  ggplot2::labs(
    color="",fill="",x="IoU of correct precitions"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 9)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
```

the majority of correct predictions had high overlap with the actual, ground truth piles with only a few piles at one site approaching the 45% minimum threshold to determine a match. let's check out the summary stats

```{r}
iou_dta_temp %>% 
  dplyr::group_by(site,method) %>% 
  dplyr::summarise(
    dplyr::across(
      c(iou)
      , .fns = list(
        mean = ~mean(.x,na.rm=T)
        , q50 = ~quantile(.x,na.rm=T,probs=0.5)
        , sd = ~sd(.x,na.rm=T)
        # , q10 = ~quantile(.x,na.rm=T,probs=0.1)
        # , q90 = ~quantile(.x,na.rm=T,probs=0.9)
        # , min = ~min(.x,na.rm=T)
        # , max = ~max(.x,na.rm=T)
        , range = ~paste0(
          scales::percent(min(.x,na.rm=T), accuracy = 0.1)
          ,"-"
          , scales::percent(max(.x,na.rm=T), accuracy = 0.1)
        )
      )
    )
    , n = dplyr::n() %>% scales::comma(accuracy = 1)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::percent(.x,accuracy=0.1)
    )
  ) %>% 
  dplyr::relocate(site,method,n) %>% 
  dplyr::arrange(site,method) %>% 
  kableExtra::kbl(
    caption = "IoU of correct predictions (TP)"
    , col.names = c(
      "site", "method"
      , "TP predictions"
      , "IoU mean"
      , "IoU median"
      , "IoU sd"
      , "IoU range"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

```{r, include=FALSE, eval=FALSE}
iou_dta_temp %>% 
  # dplyr::glimpse()
  ggplot2::ggplot(mapping = ggplot2::aes(x = iou, color = match_grp, fill = match_grp)) +
  ggplot2::geom_vline(xintercept = 0.45, color = "gray", linetype = "dashed", lwd = 1.5) +
  ggplot2::geom_density(mapping = ggplot2::aes(y=ggplot2::after_stat(scaled)), alpha = 0.8) +
  ggplot2::scale_color_manual(values=pal_match_grp) +
  ggplot2::scale_fill_manual(values=pal_match_grp) +
  ggplot2::scale_y_continuous(NULL,breaks=NULL) +
  ggplot2::scale_x_continuous(limits = c(0,1), labels=scales::percent) +
  ggplot2::facet_grid(
    rows = dplyr::vars(site)
    , cols = dplyr::vars(method)
    # , scales = "free_y"
    , switch = "y"
    , axes = "all_x"
  ) +
  ggplot2::labs(
    color="",fill="",x="IoU of correct precitions"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 11)
    , axis.text.y = ggplot2::element_blank()
    , axis.ticks.y = ggplot2::element_blank()
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

let's look at the results spatially noting that just because a prediction overlaps with a ground truth pile does not mean that it meets the IoU threshold for determining a positive match. in these cases where the IoU was insufficient, a commission (FP) will overlap with an omission (FN).

```{r, results=F}
plt_fn_temp <- function(x, method) {
  if(tolower(method) == "dbscan"){
    preds <- dbscan_spectral_preds
    gt_pred_match <- dbscan_gt_pred_match
  }else if(tolower(method) == "watershed"){
    preds <- watershed_spectral_preds
    gt_pred_match <- watershed_gt_pred_match
  }else{return(NULL)}
  # plot it
    ggplot2::ggplot() +
      ggplot2::geom_sf(
        data = stand_boundary[[x]] %>% 
          sf::st_transform(sf::st_crs(preds[[x]]))
        , color = "black", fill = NA
      ) + 
      ggplot2::geom_sf(
        data = 
          slash_piles_polys[[x]] %>% 
            dplyr::filter(is_in_stand) %>% 
            dplyr::left_join(
              gt_pred_match[[x]] %>% 
                dplyr::select(pile_id,match_grp)
              , by = "pile_id"
            ) %>% 
            sf::st_transform(sf::st_crs(preds[[x]]))
        , mapping = ggplot2::aes(fill = match_grp)
        , color = NA ,alpha=0.7
        , show.legend = T
      ) + 
      ggplot2::geom_sf(
        data =
          preds[[x]] %>%
            dplyr::filter(is_in_stand) %>% 
            dplyr::left_join(
              gt_pred_match[[x]] %>% 
                dplyr::select(pred_id,match_grp)
              , by = "pred_id"
            )
        , mapping = ggplot2::aes(fill = match_grp, color = match_grp)
        , alpha = 0
        , lwd = 0.5
        , show.legend = T
      ) +
      ggplot2::scale_fill_manual(values = pal_match_grp, drop = F, name = "") +
      ggplot2::scale_color_manual(values = pal_match_grp, drop = F, name = "") +
      ggplot2::labs(
        subtitle = paste0(
          method
          , " predictions: "
          , slash_piles_polys[[x]] %>% 
            dplyr::slice(1) %>% 
            dplyr::pull(site)
        )
      ) +
      ggplot2::theme_void() +
      ggplot2::theme(
        plot.subtitle = ggplot2::element_text(
          size = 7, hjust = 0.5
          , vjust = 1
          , margin = margin(t = 0, r = 0, b = -2, l = 0, unit = "pt")
        )
      ) +
      ggplot2::guides(
        fill = ggplot2::guide_legend(
          override.aes = list(
            color = c(NA,pal_match_grp["commission"],NA)
            , fill = c(pal_match_grp["omission"],NA,pal_match_grp["true positive"])
          )
        )
        , color = "none"
      )
}
# do it
dbscan_plt_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    \(xx)
    plt_fn_temp(xx,method = "DBSCAN")
  )
# dbscan_plt_temp[[1]]

watershed_plt_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    \(xx)
    plt_fn_temp(xx,method = "Watershed")
  )
# watershed_plt_temp[[1]]

```

combine plots with `patchwork`

```{r, fig.height=10.5, fig.width=7}
# c(dbscan_plt_temp, watershed_plt_temp) %>%
#   (\(x) x[order(names(x))])() %>% 
#   names()
patchwork::wrap_plots(
  c(dbscan_plt_temp, watershed_plt_temp) %>%
    (\(x) x[order(names(x))])()
  , ncol = 2, guides = "collect"
  , byrow = T
) &
ggplot2::theme(legend.position = "bottom")
```

note the color legend...we like to see a lot of yellow :D

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Detection Accuracy Metrics

now we aggregate the pile-level data (TP, FP, FN) into a single record for each study site and segmentation method combination. Detection performance metrics include F-score, precision, and recall.

### Aggregate Instance Match Results

because our `agg_ground_truth_match()` function that we [defined earlier](#quant_metrics_form) includes capabilities to compute quantification metrics as well (e.g. height MAPE, RMSE, etc.), we'll add the pile sizing metrics to our instance match data so that `agg_ground_truth_match()` recognizes them.

* "diff_" columns are calculated as the predicted value minus the actual value (e.g. `pred_diameter_m - gt_diameter_m`) 
* "pct_diff_" columns are calculated as the actual value minus the predicted value divided by the actual value (e.g. `(gt_diameter_m - pred_diameter_m)/gt_diameter_m`)

```{r}
# add prediction and validation measurements for dbscan
dbscan_gt_pred_match <- 
  dbscan_gt_pred_match %>% 
  purrr::imap(
    \(x,nm)
    # because psinf has field data but no others do
    if(nm=="psinf"){
      x %>% 
      # join on gt data
      dplyr::left_join(
        slash_piles_polys[[nm]] %>% 
          dplyr::rename(field_volume_m3=field_gt_volume_m3) %>% 
          dplyr::select(pile_id, image_gt_diameter_m, image_gt_area_m2, field_height_m, field_diameter_m, field_volume_m3) %>% 
          dplyr::rename_with(~ stringr::str_remove(.x, "^image_")) %>% 
          sf::st_drop_geometry()
        , by = "pile_id"
      ) %>% 
      # join on pred area data
      dplyr::left_join(
        dbscan_spectral_preds[[nm]] %>% 
          dplyr::select(
            pred_id
            , diameter_m
            , area_m2
            , volume_m3
            , max_height_m
          ) %>% 
          dplyr::rename(height_m=max_height_m) %>% 
          dplyr::rename_with(
            ~ paste0("pred_", .x, recycle0 = TRUE)
            , .cols = -c(pred_id)
          ) %>% 
          sf::st_drop_geometry()
        , by = "pred_id"
      ) %>% 
      # calculate difference columns
      dplyr::mutate(
        # area_m2
        diff_area_m2 = pred_area_m2-gt_area_m2
        , pct_diff_area_m2 = (gt_area_m2-pred_area_m2)/gt_area_m2
        # diameter_m
        , diff_diameter_m = pred_diameter_m-gt_diameter_m
        , pct_diff_diameter_m = (gt_diameter_m-pred_diameter_m)/gt_diameter_m
        # field diameter_m
        , diff_field_diameter_m = pred_diameter_m-field_diameter_m
        , pct_diff_field_diameter_m = (field_diameter_m-pred_diameter_m)/field_diameter_m
        # field height_m
        , diff_field_height_m = pred_height_m-field_height_m
        , pct_diff_field_height_m = (field_height_m-pred_height_m)/field_height_m
        # # field volume_m3
        # , diff_field_volume_m3 = pred_volume_m3-field_volume_m3
        # , pct_diff_field_volume_m3 = (field_volume_m3-pred_volume_m3)/field_volume_m3
      )  
    }else{
      x %>% 
      # join on gt data
      dplyr::left_join(
        slash_piles_polys[[nm]] %>% 
          dplyr::select(pile_id, image_gt_diameter_m, image_gt_area_m2) %>% 
          dplyr::rename_with(~ stringr::str_remove(.x, "^image_")) %>% 
          sf::st_drop_geometry()
        , by = "pile_id"
      ) %>% 
      # join on pred area data
      dplyr::left_join(
        dbscan_spectral_preds[[nm]] %>% 
          dplyr::select(
            pred_id
            , diameter_m
            , area_m2
            , volume_m3
            , max_height_m
          ) %>% 
          dplyr::rename(height_m=max_height_m) %>% 
          dplyr::rename_with(
            ~ paste0("pred_", .x, recycle0 = TRUE)
            , .cols = -c(pred_id)
          ) %>% 
          sf::st_drop_geometry()
        , by = "pred_id"
      ) %>% 
      # calculate difference columns
      dplyr::mutate(
        # area_m2
        diff_area_m2 = pred_area_m2-gt_area_m2
        , pct_diff_area_m2 = (gt_area_m2-pred_area_m2)/gt_area_m2
        # diameter_m
        , diff_diameter_m = pred_diameter_m-gt_diameter_m
        , pct_diff_diameter_m = (gt_diameter_m-pred_diameter_m)/gt_diameter_m
      )
    }
  )
# dbscan_gt_pred_match %>% dplyr::glimpse()
# add prediction and validation measurements for watershed
watershed_gt_pred_match <- 
  watershed_gt_pred_match %>% 
  purrr::imap(
    \(x,nm)
    # because psinf has field data but no others do
    if(nm=="psinf"){
      x %>% 
      # join on gt data
      dplyr::left_join(
        slash_piles_polys[[nm]] %>% 
          dplyr::rename(field_volume_m3=field_gt_volume_m3) %>% 
          dplyr::select(pile_id, image_gt_diameter_m, image_gt_area_m2, field_height_m, field_diameter_m, field_volume_m3) %>% 
          dplyr::rename_with(~ stringr::str_remove(.x, "^image_")) %>% 
          sf::st_drop_geometry()
        , by = "pile_id"
      ) %>% 
      # join on pred area data
      dplyr::left_join(
        watershed_spectral_preds[[nm]] %>% 
          dplyr::select(
            pred_id
            , diameter_m
            , area_m2
            , volume_m3
            , max_height_m
          ) %>% 
          dplyr::rename(height_m=max_height_m) %>% 
          dplyr::rename_with(
            ~ paste0("pred_", .x, recycle0 = TRUE)
            , .cols = -c(pred_id)
          ) %>% 
          sf::st_drop_geometry()
        , by = "pred_id"
      ) %>% 
      # calculate difference columns
      dplyr::mutate(
        # area_m2
        diff_area_m2 = pred_area_m2-gt_area_m2
        , pct_diff_area_m2 = (gt_area_m2-pred_area_m2)/gt_area_m2
        # diameter_m
        , diff_diameter_m = pred_diameter_m-gt_diameter_m
        , pct_diff_diameter_m = (gt_diameter_m-pred_diameter_m)/gt_diameter_m
        # field diameter_m
        , diff_field_diameter_m = pred_diameter_m-field_diameter_m
        , pct_diff_field_diameter_m = (field_diameter_m-pred_diameter_m)/field_diameter_m
        # field height_m
        , diff_field_height_m = pred_height_m-field_height_m
        , pct_diff_field_height_m = (field_height_m-pred_height_m)/field_height_m
        # # field volume_m3
        # , diff_field_volume_m3 = pred_volume_m3-field_volume_m3
        # , pct_diff_field_volume_m3 = (field_volume_m3-pred_volume_m3)/field_volume_m3
      )  
    }else{
      x %>% 
      # join on gt data
      dplyr::left_join(
        slash_piles_polys[[nm]] %>% 
          dplyr::select(pile_id, image_gt_diameter_m, image_gt_area_m2) %>% 
          dplyr::rename_with(~ stringr::str_remove(.x, "^image_")) %>% 
          sf::st_drop_geometry()
        , by = "pile_id"
      ) %>% 
      # join on pred area data
      dplyr::left_join(
        watershed_spectral_preds[[nm]] %>% 
          dplyr::select(
            pred_id
            , diameter_m
            , area_m2
            , volume_m3
            , max_height_m
          ) %>% 
          dplyr::rename(height_m=max_height_m) %>% 
          dplyr::rename_with(
            ~ paste0("pred_", .x, recycle0 = TRUE)
            , .cols = -c(pred_id)
          ) %>% 
          sf::st_drop_geometry()
        , by = "pred_id"
      ) %>% 
      # calculate difference columns
      dplyr::mutate(
        # area_m2
        diff_area_m2 = pred_area_m2-gt_area_m2
        , pct_diff_area_m2 = (gt_area_m2-pred_area_m2)/gt_area_m2
        # diameter_m
        , diff_diameter_m = pred_diameter_m-gt_diameter_m
        , pct_diff_diameter_m = (gt_diameter_m-pred_diameter_m)/gt_diameter_m
      )
    }
  )
# watershed_gt_pred_match %>% dplyr::glimpse()
# agg_ground_truth_match(watershed_gt_pred_match[[1]]) %>% dplyr::glimpse()
```

apply our `agg_ground_truth_match()` function

```{r}
agg_ground_truth_match_ans <- 
  dplyr::bind_rows(
    dbscan_gt_pred_match %>% 
      purrr::map_dfr(agg_ground_truth_match, .id = "site_data_lab") %>% 
      dplyr::mutate(method = "dbscan") %>% 
      dplyr::bind_rows()
    , watershed_gt_pred_match %>% 
      purrr::map_dfr(agg_ground_truth_match, .id = "site_data_lab") %>% 
      dplyr::mutate(method = "watershed") %>% 
      dplyr::bind_rows()
  ) %>% 
  dplyr::inner_join(
    all_stand_boundary %>% 
      dplyr::select(site_data_lab, site) %>% 
      sf::st_drop_geometry()
    , by = "site_data_lab"
  ) %>% 
  # make factor
  dplyr::mutate(
    site = ordered(site)
    , method = method %>%
      factor(ordered = T) %>% 
      forcats::fct_recode("DBSCAN" = "dbscan", "Watershed" = "watershed")
  )
```

what did we get?

```{r}
agg_ground_truth_match_ans %>% 
  dplyr::glimpse()
```

### Detection Accuracy Results

let's make a nice table of the detection accuracy metrics by study site and segmentation method

```{r}
agg_ground_truth_match_ans %>% 
  dplyr::select(
    site, method
    , tidyselect::ends_with("_n")
    , precision, recall, f_score
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      tidyselect::ends_with("_n")
      , ~scales::comma(.x,accuracy=1)
    )
    , dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::percent(.x,accuracy=0.1)
    )
  ) %>% 
  dplyr::arrange(site,method) %>% 
  kableExtra::kbl(
    caption = "Detection Accuracy"
    , col.names = c(
      "site", "method"
      , "TP predictions", "FP predictions", "FN predictions"
      , "Precision", "Recall", "F-score"
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")  
```

## Segmentation Method Accuracy Comparison

Our slash pile detection methodology implements and tests two different segmentation approaches to identify candidate slash piles from the CHM. We fully reviewed these two approaches [here](#seg_methods) but we'll present a quick review. First, the watershed segmentation method is a raster-based technique that treats the inverted height surface as a topographic landscape where local maxima serve as initial seeds to define the boundaries of individual basins. Second, DBSCAN segmentation is traditionally a point-based clustering method that we adapted for raster data by treating the cell centroids as points. To maintain consistency across varying datasets and limit the amount of effort required by users, we developed a dynamic parameterization logic that automatically calibrates both algorithms by translating user-defined physical dimensions into data-specific values. This dynamic parameterization uses the target pile size and input data resolution to calculate geometry-based search windows and density thresholds to ensure that the algorithm settings are proportional to the target object regardless of changes in input data resolution.

let's make some plots to visualize this information

```{r}
agg_ground_truth_match_ans %>% 
  dplyr::select(
    site, method
    , precision, recall, f_score
  ) %>% 
  tidyr::pivot_longer(
    cols = c(precision,recall,f_score)
    , names_to = "metric"
    , values_to = "value"
  ) %>% 
  dplyr::mutate(
    metric = metric %>%
      forcats::fct_relevel("f_score", "recall", "precision") %>%
      forcats::fct_recode("F-score" = "f_score", "Recall" = "recall", "Precision" = "precision")
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::where(is.numeric)
      , list(lab = ~scales::percent(.x,accuracy=0.1))
    )
  ) %>% 
  # dplyr::pull(metric) %>% levels()
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      y = value, x = method
      , color = metric, group = metric
    )
  ) +
  ggplot2::geom_line(alpha = 0.8, lwd = 2) +
   # lhs labs
  ggrepel::geom_text_repel(
    data = function(x){dplyr::filter(x,method == "DBSCAN")}
    , mapping = ggplot2::aes(label = value_lab, fontface = "bold")
    # , color = "black"
    , size = 3
    # , min.segment.length = 0
    , nudge_x = -0.1
    , direction = "y"
    , hjust = 1
  ) +
  # rhs labs
  ggrepel::geom_text_repel(
    data = function(x){dplyr::filter(x,method == "Watershed")}
    , mapping = ggplot2::aes(label = value_lab, fontface = "bold")
    # , color = "black"
    , size = 3
    # , min.segment.length = 0
    , nudge_x = 0.1
    , direction = "y"
    , hjust = 0
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(values=pal_eval_metric2) +
  ggplot2::scale_y_continuous(
    limits = c(0,1)
    , labels=scales::percent
    , breaks = scales::breaks_extended(n=6)
    , expand = ggplot2::expansion(mult = c(0,0.03))
  ) +
  ggplot2::facet_grid(
    cols = dplyr::vars(site)
    , axes = "all"
  ) +
  ggplot2::labs(
    color="",x="Segmentation Method",y=""
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
    , strip.text = ggplot2::element_text(size = 10, color = "black", face = "bold")
    , axis.text.x = ggplot2::element_text(size = 10)
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(override.aes = list(shape = 15, linetype = 0, size = 6, alpha = 1))
  )
```

```{r, include=F, eval = F}
# radar plot?
agg_ground_truth_match_ans %>% 
  dplyr::select(
    site, method
    , precision, recall, f_score
  ) %>% 
  tidyr::pivot_longer(
    cols = c(precision,recall,f_score)
    , names_to = "metric"
    , values_to = "value"
  ) %>% 
  dplyr::mutate(
    metric = metric %>%
      forcats::fct_relevel("f_score", "recall", "precision") %>%
      forcats::fct_recode("F-score" = "f_score", "Recall" = "recall", "Precision" = "precision")
    , site_sht = stringr::word(site) %>% ordered()
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::where(is.numeric)
      , list(lab = ~scales::percent(.x,accuracy=0.1))
    )
  ) %>% 
  # radar plot? 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = site_sht, y = value, group = method, color = method, fill = method)
  ) +
  ggplot2::geom_area(alpha = 0.5, position = "identity") +
  ggplot2::geom_point(
    data = agg_temp
    , size = 3
  ) +
  ggplot2::coord_radial(clip = "off") + 
  ggplot2::scale_y_continuous(limits=c(0,1)) +
  ggplot2::facet_grid(
    cols = dplyr::vars(metric)
    # , axes = "all"
  ) +
  ggplot2::theme_minimal()
  

# # Reshape your data for ggradar
# df_wide <- df %>%
#   pivot_wider(names_from = my_factor, values_from = value) %>%
#   rename(group = my_other_factor) # ggradar looks for a 'group' column
# 
# # Create the plot
# ggradar(df_wide, 
#         values.radar = c("0", "0.5", "1"), # Labels for the grid
#         grid.min = 0, grid.mid = 0.5, grid.max = 1,
#         fill = TRUE,         # This enables the origin-to-point fill
#         fill.alpha = 0.3) +  # Transparency for overlapping areas
#   theme(legend.position = "bottom")  
```

recall never changes, the changes in F-score are entirely due to changes in precision

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

### Paired T-Test

To compare the detection performance of the two algorithms across our set of four study sites, we can quickly use a paired t-test to check if the average of the differences in F-score between the two methods is significantly different from zero. If the p-value is less than 0.05, you can say the difference is statistically significant. However, with only four sites this test may not be robust enough to detect differences in detection accuracy across the methods unless one method is a lot better...

```{r}
# get scores for each method since t.test needs the sites to be ORDERED the same for proper "paired" comparison
dbscan_temp <- agg_ground_truth_match_ans %>% 
  dplyr::arrange(site) %>% 
  dplyr::filter(tolower(method) == "dbscan") %>% 
  dplyr::pull(f_score)
watershed_temp <- agg_ground_truth_match_ans %>% 
  dplyr::arrange(site) %>% 
  dplyr::filter(tolower(method) == "watershed") %>% 
  dplyr::pull(f_score)
# t.test
ttest_temp <- t.test(dbscan_temp, watershed_temp, paired = T)
```

let's see what we got

```{r}
ttest_temp
```

the `t.test()` computed the mean difference (MD) between the DBSCAN and Watershed F-score to determine if the difference is statistically significant (i.e. significantly different from zero). the mean difference (MD) is `r scales::percent(ttest_temp$estimate, accuracy = 0.01)` (DBSCAN F-Score minus Watershed F-Score). the p-value of `r scales::comma(ifelse(ttest_temp$p.value<0.00001,0.00001,ttest_temp$p.value), accuracy = 0.0001)` is greater than 0.05, meaning we fail to reject the null hypothesis that the true mean difference is zero. this implies that neither the DBSCAN nor the Watershed method was statistically significantly better (or worse) than the other segmentation method at correctly identifying the location of actual slash piles while minimizing incorrect predictions at locations where no actual pile was present.

### Bayesian Approach

The paired t-test above failed to detect a difference in detection accuracy (F-score) by segmentation method (DBSCAN vs Watershed). With only four sites the paired t-test may not be robust enough to detect differences in detection accuracy across the methods. An alternative is a quick Bayesian model which may provide a more robust alternative to the paired t-test by estimating the full probability distribution of the difference between methods rather than relying on a single p-value. While a paired t-test simply evaluates the average difference within sites, a Bayesian model can account for the nested structure of the data by allowing the intercept to vary by study site (`(1|site)` in the model) while keeping the slope constant across locations (`f_score ~ method` in the model). The slope of the model represents the effect of the segmentation method. This approach allows the model to adjust for the unique baseline difficulty of pile detection at each study site and isolates the detection accuracy change caused by using a different segmentation method. Since F-score is bound between zero and one, we use a Beta regression family to ensure predictions are not made outside of these bounds.

```{r}
# brms::brm
fscore_mod_method_temp <- brms::brm(
  f_score ~ method + (1|site)
  , data = agg_ground_truth_match_ans
  # family
  , family = brms::Beta(link = "logit")
  # mcmc
  , iter = 6000, warmup = 3000
  , chains = 4
  , cores = lasR::half_cores()
  , file = paste0("../data/", "fscore_mod_method")
)
# look at results
# plot(fscore_mod_method_temp)
```

quickly look at the model estimates of the effect of segmentation method

```{r}
brms::fixef(fscore_mod_method_temp)
# stats::coef(fscore_mod_method_temp)
```

plot the posterior predictive distributions of the conditional mean F-score by segmentation method with the 95% highest posterior density interval (HDI)

```{r}
# get posterior draws
draws_temp <- agg_ground_truth_match_ans %>%
  dplyr::distinct(method) %>% 
  # re_formula = NA gives population average (average across site)
  tidybayes::add_epred_draws(fscore_mod_method_temp, re_formula = NA) %>% 
  dplyr::mutate(value = .epred) %>% 
  dplyr::ungroup()
# plot
draws_temp %>% 
  ggplot(
    mapping = aes(
      x = value
      , y = method
      , fill = method
    )
  ) +
  tidybayes::stat_halfeye(
    point_interval = tidybayes::median_hdci, .width = .95
    , interval_color = "gray66"
    , shape = 21, point_color = "gray66", point_fill = "black"
    , justification = -0.01
  ) +
  harrypotter::scale_fill_hp_d(option = "dracomalfoy", drop = F) +
  ggplot2::scale_x_continuous(limits = c(0,1), labels=scales::percent, breaks = scales::breaks_extended(n=6)) +
  ggplot2::labs(
    subtitle = "posterior distribution of F-score by segmentation method\nand the 95% highest posterior density interval (HDI)"
    , x = "F-score", y = ""
  ) +
  theme_light() +
  theme(legend.position = "none")
```

summary table of the posterior distribution and 95% HDI using `tidybayes::median_hdci()` to avoid potential for returning multiple rows by group if our data is grouped. See the [documentation](https://mjskay.github.io/ggdist/reference/point_interval.html) for the `ggdist` package which notes that "If the distribution is multimodal, hdi may return multiple intervals for each probability level (these will be spread over rows)."

```{r}
draws_temp %>%
  dplyr::group_by(method) %>% 
  tidybayes::median_hdci(value) %>% 
  dplyr::select(-c(.point,.interval, .width)) %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::percent(.x,accuracy=0.1)
    )
  ) %>% 
  kableExtra::kbl(
    digits = 2
    , caption = "F-score<br>95% HDI of the posterior predictive distribution"
    , col.names = c(
      "method"
      , "F-score<br>median"
      , "HDI low", "HDI high"
    )
    , escape = F
  ) %>% 
  kableExtra::kable_styling()
```

we can also make pairwise comparisons using the posterior draws with `tidybayes::compare_levels()` to determine the difference in F-score between the two methods, but now, instead of one mean difference value as in the paired t-test, we get a distribution of the difference which provides much more detailed insight to the difference in detection accuracy between the two segmentation methods

```{r}
comp_temp <- draws_temp %>%
  tidybayes::compare_levels(
    value
    , by = method
    # make the comparison the same as t.test: dbscan-watershed
    , comparison = list(c("DBSCAN","Watershed"))
  )
# dplyr::glimpse(comp_temp)
med_comp_temp <- tidybayes::median_hdci(comp_temp$value)
mean_comp_temp <- tidybayes::mean_hdci(comp_temp$value)
# plot
comp_temp %>% 
  ggplot2::ggplot(aes(x = value, y = 0)) +
  ggplot2::geom_vline(
    xintercept = 0
    # , linetype = "dashed"
    , color = "black"
    , lwd = 1.5
  ) +
  tidybayes::stat_halfeye(
    point_interval = tidybayes::median_hdci, .width = .95
    , fill = "navy", alpha = 0.8
    , interval_color = "gray66"
    , shape = 21, point_color = "gray66", point_fill = "black"
    , justification = -0.01
  ) +
  ggplot2::annotate(
    "text"
    , x = (med_comp_temp %>% dplyr::select(y,ymin,ymax) %>% base::unlist(use.names = F))
    , y = -0.01
    , label = (
      med_comp_temp %>% 
        dplyr::select(y,ymin,ymax) %>% 
        base::unlist(use.names = F) %>% 
        scales::percent(accuracy = 0.1)
    )
  ) +
  ggplot2::scale_y_continuous(breaks=NULL) +
  ggplot2::scale_x_continuous(labels=scales::percent, breaks = scales::breaks_extended(n=8), expand = ggplot2::expansion(mult = 0.05)) +
  ggplot2::labs(
    subtitle = "difference in F-score between segmentation methods"
    , x = "difference in F-score"
    , y = "difference (DBSCAN - Watershed)"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    axis.title.y = ggplot2::element_text(angle = 0, vjust = 0.5)
  )
# scales::percent(mean_comp_temp$y,accuracy=0.1)
# scales::percent(ttest_temp$estimate, accuracy = 0.1)
```

the 95% HDI shown as the gray line below the posterior distribution summarizes the posterior draws to represent a density interval with a value of central tendency. We plot the median (`scales::percent(med_comp_temp$y,accuracy=0.1)`) as the value of central tendency of the difference in F-score between the two methods but we could also determine the mean difference which we calculated above using `tidybayes::mean_hdci()`. Our Bayesian approach estimates the mean difference in F-score between the two methods as `scales::percent(mean_comp_temp$y,accuracy=0.1)` which we can directly compare to the mean difference tested using the paired t-test above of `scales::percent(ttest_temp$estimate, accuracy = 0.1)`.

table the 95% HDI of the difference in F-score between the two methods

```{r}
med_comp_temp %>% 
  dplyr::mutate(
    dplyr::across(
      dplyr::where(is.numeric)
      , ~scales::percent(.x,accuracy=0.1)
    )
  ) %>% 
  dplyr::select(tidyselect::starts_with("y")) %>% 
  dplyr::mutate(d = "difference (DBSCAN - Watershed)") %>% 
  dplyr::relocate(d) %>% 
  kableExtra::kbl(
    caption = "difference in F-score between the two methods<br>95% HDI of the posterior predictive distribution"
    , col.names = c(
      ""
      , "difference F-score<br>median"
      , "HDI low", "HDI high"
    )
    , escape = F
  ) %>% 
  kableExtra::kable_styling()
```

finally, we can use the posterior draws of the difference in F-score (calculated with `tidybayes::compare_levels()`) to determine the probability that one method is better than the other

```{r}
comp_temp <- comp_temp %>% 
  dplyr::mutate(
    diff_gt_0 = as.numeric(value>0)
  ) 
# simply calculate the proportion of posterior draws where the difference is greater than zero
mean(comp_temp$diff_gt_0) %>% scales::percent(accuracy = 0.1)
```

There is a `scales::percent(mean(comp_temp$diff_gt_0),accuracy = 0.1)` probability that the DBSCAN method results in better slash pile detection accuracy (i.e. higher F-score) than the Watershed method

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Individual Pile Evaluation

let's visually inspect the TP, FP, and FN predictions to see if we can get some insight into why actual piles were missed (omissions) or correctly predicted (TP) and why the method may have made incorrect predictions (commissions)

```{r}
p_fn_temp <- function(row_n = 1, site, method = "dbscan", mtchgrp = "true positive") {
  if(tolower(method) == "dbscan"){
    w_dta <- dbscan_gt_pred_match[[site]] %>% 
      dplyr::filter(match_grp == mtchgrp)
    p_dta <- dbscan_spectral_preds[[site]]
  }else{
    w_dta <- watershed_gt_pred_match[[site]] %>% 
      dplyr::filter(match_grp == mtchgrp)
    p_dta <- watershed_spectral_preds[[site]]
  }
  # get area of largest with buffer so all plots are same size
  if(site %in% c("bhef")){
    b <- 8.5
  }else if(site %in% c("arnf")){
    b <- 6
  }else{
    b <- 4
  }
  
  if(
    mtchgrp %in%  c("true positive", "omission")
  ){
    area_temp <- slash_piles_polys[[site]] %>% 
      dplyr::inner_join(
        w_dta %>% 
          dplyr::slice_max(order_by = gt_area_m2, n = 1) 
        , by = "pile_id"
      ) %>% 
      sf::st_transform(crs=terra::crs(rgb_rast[[site]])) %>% 
      sf::st_buffer(b) %>% 
      sf::st_area() %>% 
      as.numeric() %>% 
      ceiling() 
  }else{
    area_temp <- p_dta %>% 
      dplyr::inner_join(
        w_dta %>% 
          dplyr::slice_max(order_by = pred_area_m2, n = 1) 
        , by = "pred_id"
      ) %>% 
      sf::st_transform(crs=terra::crs(rgb_rast[[site]])) %>% 
      sf::st_buffer(b) %>% 
      sf::st_area() %>% 
      as.numeric() %>% 
      ceiling()
  }
  # tp
  tp <- w_dta %>% dplyr::slice(row_n) 
  # gt pile
  gt_pile <- slash_piles_polys[[site]] %>% 
    dplyr::filter(pile_id == tp$pile_id)
  # pred pile
  pred_pile <- p_dta %>% 
    dplyr::filter(pred_id == tp$pred_id)
  
  # ortho area
  if(nrow(gt_pile) == 1 && nrow(pred_pile) == 1){
    ortho_st <- 
      sf::st_union(
        gt_pile %>% sf::st_transform(crs=terra::crs(rgb_rast[[site]])) 
        , pred_pile %>% sf::st_transform(crs=terra::crs(rgb_rast[[site]])) 
      ) %>% 
      sf::st_centroid() %>% 
      sf::st_buffer(
        sqrt(area_temp/4) ## numerator = desired plot size in m2
        , endCapStyle = "SQUARE"
      ) %>%
      dplyr::mutate(dummy=1)  
  }else if(nrow(gt_pile) == 1){
    ortho_st <- 
      gt_pile %>% 
      sf::st_transform(crs=terra::crs(rgb_rast[[site]])) %>% 
      sf::st_centroid() %>% 
      sf::st_buffer(
        sqrt(area_temp/4) ## numerator = desired plot size in m2
        , endCapStyle = "SQUARE"
      ) %>%
      dplyr::mutate(dummy=1)  
  }else if(nrow(pred_pile) == 1){
    ortho_st <- 
      pred_pile %>% 
      sf::st_transform(crs=terra::crs(rgb_rast[[site]])) %>% 
      sf::st_centroid() %>% 
      sf::st_buffer(
        sqrt(area_temp/4) ## numerator = desired plot size in m2
        , endCapStyle = "SQUARE"
      ) %>%
      dplyr::mutate(dummy=1)  
  }
  
  # get plt
  rp <- ortho_plt_fn(
    rgb_rast = rgb_rast[[site]]
    , stand = ortho_st
    , add_stand = F
    , buffer = 0
  ) 
  
  # #### add chm?
  # if(addchm){
  #   # get max ht
  #   maxh <- all_stand_boundary %>% 
  #     dplyr::filter(site_data_lab == site) %>% 
  #     dplyr::slice(1) %>% 
  #     dplyr::pull(max_ht_m)
  #   # ff <- tempfile(fileext = ".tif")
  #   # crop, slice rast
  #   ##### ???????????????????????? this kills your computer every time :\
  #   crp_chm_rast_temp <- chm_rast[[site]] %>% 
  #     terra::crop(
  #       ortho_st %>% 
  #         dplyr::ungroup() %>% 
  #         sf::st_union() %>% 
  #         sf::st_bbox() %>% 
  #         sf::st_as_sfc() %>% 
  #         sf::st_transform(terra::crs(chm_rast[[site]])) %>% 
  #         terra::vect()
  #       , mask = T
  #       # , filename = ff
  #     )
  #   crp_chm_rast_temp <- 
  #     # slice
  #     terra::clamp(
  #       crp_chm_rast_temp
  #       , upper = maxh
  #       , lower = 0
  #       , values = F
  #     )
  #   rp <- rp +
  #     ggnewscale::new_scale_fill() +
  #     ggplot2::geom_tile(
  #       data = crp_chm_rast_temp %>%
  #         terra::as.data.frame(xy=T) %>%
  #         dplyr::rename(f=3)
  #       , mapping = ggplot2::aes(x=x,y=y,fill=f)
  #       , alpha = 0.5
  #       , inherit.aes = F
  #     ) +
  #     ggplot2::scale_fill_viridis_c(
  #       option = "plasma", na.value = "gray",name = "CHM (m)"
  #       , limits = c(0,ceiling(maxh))
  #     )
  # }
  
  #### add piles
  if(nrow(gt_pile) == 1){
    rp <- rp +
      ggplot2::geom_sf(data = gt_pile, fill = NA, color = "cyan", lwd = 0.6)
  }
  if(nrow(pred_pile) == 1){
    rp <- rp +
      ggplot2::geom_sf(data = pred_pile, fill = NA, color = "magenta", lwd = 0.5)
  }
  
  
  
  return(
    rp
    # list(rp,ortho_st)
  )
}
```

```{r,include=FALSE,eval=FALSE}
xxx <- p_fn_temp(row_n = 2, site = "pj", method = "dbscan", mtchgrp = "commission")
xxx[[1]]
# xxx
# get max ht
    (maxh <- all_stand_boundary %>% 
      dplyr::filter(site_data_lab == "pj") %>% 
      dplyr::slice(1) %>% 
      dplyr::pull(max_ht_m))
    # crop, slice rast
    crp_chm_rast_temp <- chm_rast[["pj"]] %>% 
      terra::crop(
          xxx[[2]] %>% 
          dplyr::ungroup() %>% 
          sf::st_union() %>% 
          sf::st_bbox() %>% 
          sf::st_as_sfc() %>% 
          sf::st_transform(terra::crs(chm_rast[["pj"]])) %>% 
          terra::vect()
      )
    
    crp_chm_rast_temp <- crp_chm_rast_temp %>% 
      # slice
      terra::clamp(
        upper = maxh
        , lower = 0
        , values = F
      )
    # terra::plot(crp_chm_rast_temp)
    xxx[[1]] +
      ggnewscale::new_scale_fill() +
      ggplot2::geom_tile(
        data = crp_chm_rast_temp %>% 
          terra::as.data.frame(xy=T) %>% 
          dplyr::rename(f=3)
        , mapping = ggplot2::aes(x=x,y=y,fill=f)
        , alpha = 0.5
        , inherit.aes = F
      ) +
      ggplot2::scale_fill_viridis_c(
        option = "plasma", na.value = "gray",name = "CHM (m)"
        , limits = c(0,ceiling(maxh))
      )

remove(xxx,maxh, crp_chm_rast_temp)
gc()

dbscan_tp_plts_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    function(x){
      nn <- 
        dbscan_gt_pred_match[[x]] %>% 
        dplyr::filter(match_grp == "true positive") %>% 
        nrow()
      pl <- 
        1:min(nn,2) %>% 
        purrr::map(
          \(r)
          p_fn_temp(row_n = r, site = x, method = "dbscan", mtchgrp = "true positive")  
        )
      
      # return(pl)
      
      pl2 <- pl %>% 
        purrr::map(
          function(wtfchm, site = x){
             # get max ht
            maxh <- all_stand_boundary %>% 
              dplyr::filter(site_data_lab == site) %>% 
              dplyr::slice(1) %>% 
              dplyr::pull(max_ht_m)
            # crop, slice rast
            crp_chm_rast_temp <- chm_rast[[site]] %>% 
              terra::crop(
                  wtfchm[[2]] %>% 
                  dplyr::ungroup() %>% 
                  sf::st_union() %>% 
                  sf::st_bbox() %>% 
                  sf::st_as_sfc() %>% 
                  sf::st_transform(terra::crs(chm_rast[[site]])) %>% 
                  terra::vect()
              )
            
            crp_chm_rast_temp <- crp_chm_rast_temp %>% 
              # slice
              terra::clamp(
                upper = maxh
                , lower = 0
                , values = F
              )
            # terra::plot(crp_chm_rast_temp)
            wtfchm[[1]] +
              ggnewscale::new_scale_fill() +
              ggplot2::geom_tile(
                data = crp_chm_rast_temp %>% 
                  terra::as.data.frame(xy=T) %>% 
                  dplyr::rename(f=3)
                , mapping = ggplot2::aes(x=x,y=y,fill=f)
                , alpha = 0.5
                , inherit.aes = F
              ) +
              ggplot2::scale_fill_viridis_c(
                option = "plasma", na.value = "gray",name = "CHM (m)"
                , limits = c(0,ceiling(maxh))
              )
          }
        )
      return(pl2)
    }
  )

```

### True Positives

Check out the plots for the TP matches

```{r,eval=F, results=F}
# max piles per site
nplts_temp <- 12
# dbscan
dbscan_tp_plts_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    function(x){
      nn <- 
        dbscan_gt_pred_match[[x]] %>% 
        dplyr::filter(match_grp == "true positive") %>% 
        nrow()
      pl <- 
        1:min(nn,nplts_temp) %>% 
        purrr::map(
          \(r)
          p_fn_temp(row_n = r, site = x, method = "dbscan", mtchgrp = "true positive")
        )
      return(pl)
    }
  )
# watershed
watershed_tp_plts_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    function(x){
      nn <- 
        watershed_gt_pred_match[[x]] %>% 
        dplyr::filter(match_grp == "true positive") %>% 
        nrow()
      pl <- 
        1:min(nn,nplts_temp) %>% 
        purrr::map(
          \(r)
          p_fn_temp(row_n = r, site = x, method = "watershed", mtchgrp = "true positive")  
        )
      return(pl)
    }
  )
```

#### DBSCAN

plot example true positive matches on the RGB with the actual pile (cyan) and predicted (magenta) pile outlined

```{r,eval=F, fig.height=8.8, results='hide', fig.show='asis'}
all_stand_boundary$site_data_lab %>% 
purrr::map(
  \(x)
  dbscan_tp_plts_temp[[x]] %>% 
  # purrr::map(1) %>% 
  # purrr::keep_at(1:4) %>% 
  patchwork::wrap_plots(ncol = 4) +
  patchwork::plot_annotation(
    title = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(site)
    , theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 10, hjust = 0.5))
  )
)
  
```

#### Watershed

plot example true positive matches on the RGB with the actual pile (cyan) and predicted (magenta) pile outlined

```{r,eval=F, fig.height=8.8, results='hide', fig.show='asis'}
all_stand_boundary$site_data_lab %>% 
purrr::map(
  \(x)
  watershed_tp_plts_temp[[x]] %>% 
  # purrr::map(1) %>% 
  patchwork::wrap_plots(ncol = 4) +
  patchwork::plot_annotation(
    title = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(site)
    , theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 10, hjust = 0.5))
  )
)
  
```

### Omissions (False Negatives)

Check out the plots for the omissions (FN)

```{r,eval=F, results=F}
# dbscan
dbscan_tp_plts_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    function(x){
      nn <- 
        dbscan_gt_pred_match[[x]] %>% 
        dplyr::filter(match_grp == "omission") %>% 
        nrow()
      pl <- 
        1:min(nn,nplts_temp) %>% 
        purrr::map(
          \(r)
          p_fn_temp(row_n = r, site = x, method = "dbscan", mtchgrp = "omission")  
        )
      return(pl)
    }
  )
# watershed
watershed_tp_plts_temp <- all_stand_boundary$site_data_lab %>% 
  purrr::set_names() %>% 
  purrr::map(
    function(x){
      nn <- 
        watershed_gt_pred_match[[x]] %>% 
        dplyr::filter(match_grp == "omission") %>% 
        nrow()
      pl <- 
        1:min(nn,nplts_temp) %>% 
        purrr::map(
          \(r)
          p_fn_temp(row_n = r, site = x, method = "watershed", mtchgrp = "omission")  
        )
      return(pl)
    }
  )
```

#### DBSCAN

plot example omissions on the RGB with the actual pile (cyan) pile outlined

```{r,eval=F, fig.height=8.8, results='hide', fig.show='asis'}
all_stand_boundary$site_data_lab %>% 
purrr::map(
  \(x)
  dbscan_tp_plts_temp[[x]] %>% 
  # purrr::map(1) %>% 
  # purrr::keep_at(1:4) %>% 
  patchwork::wrap_plots(ncol = 4) +
  patchwork::plot_annotation(
    title = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(site)
    , theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 10, hjust = 0.5))
  )
)
  
```

why are these missing?

```{r}
# function to plot ortho
terra_plt_ortho <- function(
  ortho
  , gt_piles = NA
  , pred_piles = NA
  , boundary
  , title
  , chm = NA
  , maxh = 99
) {
  terra::plotRGB(ortho, stretch = "lin")
  terra::plot(
    boundary %>% 
      sf::st_transform(terra::crs(ortho)) %>% 
      terra::vect()
    , add = T, border = "black", col = NA
    , main = title
  )
  if(inherits(chm,"SpatRaster")){
    terra::plot(
      chm
      , add = T, col = viridis::plasma(n=100), alpha = 0.5
      , legend = F
      , range = c(0,maxh), fill_range = T
    )
  }
  if(!all(is.na(gt_piles))){
    terra::plot(
      gt_piles %>% 
        dplyr::filter(is_in_stand) %>% 
        sf::st_transform(terra::crs(ortho)) %>% 
        terra::vect()
      , add = T, border = "cyan", col = NA, lwd = 1.2
    )
  }
  if(!all(is.na(pred_piles))){
    terra::plot(
      pred_piles %>% 
        sf::st_transform(terra::crs(ortho)) %>% 
        terra::vect()
      , add = T, border = "magenta", col = NA, lwd = 1.2
    )
  }
}
# map over it 
all_stand_boundary$site_data_lab %>% 
  purrr::map(
    function(x){
    piles <- slash_piles_polys[[x]] %>% 
      dplyr::filter(is_in_stand) %>% 
      dplyr::inner_join(
        dbscan_gt_pred_match[[x]] %>% 
        dplyr::filter(match_grp == "omission") %>% 
        dplyr::select(pile_id)
        , by = "pile_id"
      )
    maxh <- all_stand_boundary %>% 
      dplyr::filter(site_data_lab == x) %>% 
      dplyr::slice(1) %>% 
      dplyr::pull(max_ht_m)
    # crop, slice rast
    crp_chm_rast_temp <- chm_rast[[x]] %>% 
      terra::crop(
          piles %>% 
          sf::st_transform(terra::crs(chm_rast[[x]])) %>% 
          terra::vect()
          , mask = T
      )
    
    crp_chm_rast_temp <- crp_chm_rast_temp %>% 
      # slice
      terra::clamp(
        upper = maxh
        , lower = 0
        , values = F
      )
    
    terra_plt_ortho(
      ortho = rgb_rast[[x]]
      , gt_piles = piles
      , pred_piles = NA
      , boundary = stand_boundary[[x]]
      , chm = crp_chm_rast_temp
      , maxh = maxh
      , title = paste0(
        "DBSCAN: "
        , all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(site)
      )
    )
  })

```


#### Watershed

plot example omissions on the RGB with the actual pile (cyan) pile outlined

```{r,eval=F, fig.height=8.8, results='hide', fig.show='asis'}
all_stand_boundary$site_data_lab %>% 
purrr::map(
  \(x)
  watershed_tp_plts_temp[[x]] %>% 
  # purrr::map(1) %>% 
  # purrr::keep_at(1:4) %>% 
  patchwork::wrap_plots(ncol = 4) +
  patchwork::plot_annotation(
    title = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(site)
    , theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 10, hjust = 0.5))
  )
)
  
```

