# Detection Accuracy{#preds_detect}

To this point we have:

1. Provided a data overview: [here](#data_desc) and [here](#data_load)
2. [Processed the UAS point cloud](#ptcld_process)
3. [Demonstrated our geometry-based slash pile detection methodology](#geom_detect)
4. [Demonstrated our spectral refinement (i.e. data fusion) methodology](#data_fusion)
5. [Reviewed how we will evaluate our method](#meth_eval)
6. [Made predictions using our method on four experimental sites](#meth_preds)
7. and [Evaluated the pile detection accuracy of the structural-plus-spectral data fusion methodology](#preds_detect)

In this section, we'll evaluate the effectiveness of the proposed geometric, rules-based slash pile detection methodology by assessing its quantification accuracy performance. We fully reviewed the quantification accuracy assessment [workflow here](#quant_metrics_form), but here is a quick overview:

While [detection accuracy](#preds_detect) evaluates the ability of the method to correctly locate slash piles, the quantification accuracy assessment measures the precision of the physical pile form dimensions estimated for the successfully identified piles. The quantification accuracy assessment focuses exclusively on TP matches to calculate performance metrics such as Root Mean Square Error (RMSE), Mean Absolute Percentage Error (MAPE), and Mean Error (ME) for pile form measurements height and diameter. Unlike the detection validation which includes all four study sites, this quantification accuracy evaluation is limited to the PSINF Mixed Conifer site where direct field-measured ground truth data is available. 

We intentionally limit the quantification accuracy reporting to measurements of height and diameter which were made directly in the field rather than derived volume. To evaluate volume, we will perform a direct comparison between the volumes calculated from our predicted segments and those calculated from field measurements using identical geometric formulas. This comparison functions as a test of measurement consistency rather than a formal accuracy assessment of the remote-sensing method itself. This approach ensures our quantification metrics reflect the actual performance of the detection method instead of a composite error involving sensor noise, field variance, and geometric assumptions. Because both the predicted and field-based volumes rely on the same simplified shape assumptions, any resulting differences are not treated as true errors. Instead, this evaluation highlights how variations in the primary inputs of height and diameter propagate through to the final volume estimate. This distinction is critical because it separates the verifiable accuracy of our sensor-derived physical pile form measurements from the subsequent modeling of three-dimensional space.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Height and Diameter Accuracy

Let's evaluate the height and diameter accuracy of our slash pile detection and quantification framework using field measurements for the PSINF site which were taken by measuring the height and diameter (longest side of pile) using a laser hypsometer

We already computed the performance metrics RMSE, MAPE, and ME for pile form measurements height and diameter at the study site and detection methodology level in the [prior section](#agg_inst_match) using our `agg_ground_truth_match()` function that we [defined earlier](#quant_metrics_form)

let's check out what we got

```{r}
df_temp <- 
  agg_ground_truth_match_ans %>% 
  dplyr::mutate(
    ref_trees = tp_n+fn_n
    , det_trees = tp_n+fp_n
  ) %>% 
  dplyr::select(
    site, method
    # , site_area_m2
    # detection
    , ref_trees
    , det_trees
    , tp_n
    , omission_rate,commission_rate,recall,precision,f_score
    # quantification
    , tidyselect::ends_with("_mean")
    , tidyselect::ends_with("_rmse")
    , tidyselect::ends_with("_mape")
  ) %>% 
  # second select to arrange pile_metric
  dplyr::select(
    site, method
    # , site_area_m2
    # detection
    , ref_trees
    , det_trees
    , tp_n
    , omission_rate,commission_rate,recall,precision,f_score
    # quantification
    # , c(tidyselect::contains("volume") & !tidyselect::contains("paraboloid"))
    , tidyselect::contains("area")
    , tidyselect::contains("height")
    , tidyselect::contains("diameter")
  ) %>% 
  # names()
  dplyr::mutate(
    dplyr::across(
      .cols = c(
        f_score, recall, precision, tidyselect::ends_with("_mape")
        , tidyselect::ends_with("_rate")
      )
      , .fn = ~ scales::percent(.x, accuracy = 1)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_mean"))
      , .fn = ~ scales::comma(.x, accuracy = 0.01)
    )
    , dplyr::across(
      .cols = c(tidyselect::ends_with("_rmse"))
      , .fn = ~ scales::comma(.x, accuracy = 0.1)
    )
  )
  # dplyr::glimpse()

  ########################## adj this if want lots of cols
df_temp %>% 
  dplyr::inner_join(
    all_stand_boundary %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(site_data_lab == "psinf") %>% 
      dplyr::select(site,site_area_ha)
    , by = "site"
  ) %>% 
  dplyr::select(
    !tidyselect::contains("_area_")
    & !tidyselect::contains("diff_diameter_")
    & !tidyselect::ends_with("_trees")
    # & !tidyselect::ends_with("_n")
    & !tidyselect::ends_with("_rate")
  ) %>% 
  dplyr::select(
    -c(recall,precision,f_score)
  ) %>% 
  ########################## adj this if want lots of cols
  # dplyr::glimpse()
  dplyr::relocate(site,method,tp_n) %>% 
  dplyr::arrange(site,method) %>% 
  kableExtra::kbl(
    caption = "Quantification Accuracy"
    , col.names = c(
      "site", "method", "TP predictions"
      , rep(c("ME","RMSE","MAPE"), times = 2)
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11.5) %>% 
  kableExtra::add_header_above(c(
    " "=3
    # , "Detection" = 3
    # , "Area" = 3
    , "Height (m)" = 3
    , "Diameter (m)" = 3
  )) %>% 
  kableExtra::column_spec(seq(3,9,by=3), border_right = TRUE, include_thead = TRUE) %>% 
  # kableExtra::column_spec(
  #   column = 3:9
  #   , extra_css = "font-size: 10px;"
  #   , include_thead = T
  # ) %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

These results demonstrate that the two segmentation methods, DBSCAN and Watershed, show no significant difference in their ability to represent the physical form of correctly identified piles. Because the quantification of pile form is derived directly from the underlying structural CHM data, the two methods naturally yield similar results once a pile is successfully detected. Since both algorithms operate on the same input data to generate measurements for a given location, major differences in quantification accuracy are only expected if there are significant disparities in detection performance. Our detection accuracy evaluation confirmed that both methods performed similarly, and as a result, they produced ~~nearly~~ identical representations of physical pile attributes.

## Area Accuracy

We could also evaluate quantification accuracy using pile areas based on the image annotations, though these are less presumptive than the "gold standard" field measurements. These annotations were occasionally limited by the difficulty of pinpointing exact pile boundaries, even when using high-resolution RGB data. Despite the potential for human error in the digitizing process, these area measurements provide additional validation data that is available across all four study sites.

```{r}
  ########################## adj this if want lots of cols
df_temp %>% 
  dplyr::select(
    !tidyselect::contains("_height_")
    & !tidyselect::contains("_diameter_")
    & !tidyselect::ends_with("_trees")
    # & !tidyselect::ends_with("_n")
    & !tidyselect::ends_with("_rate")
  ) %>% 
  dplyr::select(
    -c(recall,precision,f_score)
  ) %>% 
  ########################## adj this if want lots of cols
  # dplyr::glimpse()
  dplyr::relocate(site,method,tp_n) %>% 
  dplyr::arrange(site,method) %>% 
  kableExtra::kbl(
    caption = "Quantification Accuracy"
    , col.names = c(
      "site", "method", "TP predictions"
      , rep(c("ME","RMSE","MAPE"), times = 1)
    )
    , escape = F
    # , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 11.5) %>% 
  kableExtra::add_header_above(c(
    " "=3
    # , "Detection" = 3
    # , "Area" = 3
    , "Area (m<sup>2</sup>)" = 3
    
  ), escape = F) %>% 
  kableExtra::column_spec(seq(3,6,by=3), border_right = TRUE, include_thead = TRUE) %>% 
  # kableExtra::column_spec(
  #   column = 3:9
  #   , extra_css = "font-size: 10px;"
  #   , include_thead = T
  # ) %>% 
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

These results further confirm that the two segmentation methods yield similar quantification accuracies for pile area, mirroring the patterns observed for height and diameter. While the absolute area accuracy naturally varies across study sites according to the average pile size, the MAPE remains decently low, ranging between `r scales::percent(min(agg_ground_truth_match_ans$pct_diff_area_m2_mape),accuracy = 0.1)` and `r scales::percent(max(agg_ground_truth_match_ans$pct_diff_area_m2_mape),accuracy = 0.1)` across all locations. These metrics suggest that both DBSCAN and Watershed are capable of delineating pile boundaries with acceptable precision.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Volume Comparison

We excluded quantification accuracy metrics for derived volume because the resulting value would not constitute a true "error". Comparing our predicted volume to a volume that was *not* directly measured, but instead calculated using a geometric assumption (like assuming a perfectly circular base and paraboloid shape) would be inappropriate. This is because any resulting difference between the prediction and the ground truth would be a blend of three inseparable factors: the error of the remote-sensing prediction method, the error in the direct field measurements (diameter/height), and the error introduced by the geometric shape assumption. Reporting such combined errors would be misleading, as it would be impossible to isolate the true performance of our remote-sensing method alone.

Instead, data involving derived values of volume based on field measurements and a shape assumption and its comparison to our irregularly shaped CHM-derived volume will be treated simply as data points for insight into the differences. Using geometric shape assumptions for estimating pile volume is the standard practice when implementing prescriptions or preparing for slash pile burning ([Hardy 1996](https://permanent.fdlp.gov/gpo45282/index.htm); [Long & Boston 2014](https://doi.org/10.5849/forsci.13-501)). This comparison will help us understand the discrepancy between our irregularly shaped CHM-derived volume and the volume calculated assuming a perfectly circular base and paraboloid shape with field-measured height and diameter. This approach will still provide valuable context about the impact of the perfectly circular base and paraboloid geometric assumptions without falsely attributing the error of the simplified model to the remote-sensing method itself.

let's do that now

* **field-measured piles** 
  - **volume** assumes a paraboloid shape, with volume calculated using the field-measured diameter (as the width) and height. we'll refer to this as "Allometric Field Volume" to indicate the field measurement is derived using a shape assumption.
* **predicted piles** 
  - **volume** calculated from the elevation profile of the irregular predicted pile footprint, without assuming a specific geometric shape. we'll refer to this as "Predicted Volume" to indicate the predicted measurement is from our CHM-based detection methodology
  
We would generally expect that the allometric field volume is larger than the predicted volume because the allometric calculation assumes a perfectly regular geometric shape (circular base and paraboloid) based on maximum field dimensions (height and diameter). this process effectively encloses the actual, irregular pile form within a simplified geometric dome which inherently neglects and sits above the actual irregularities and voids in the pile structure, likely leading to an overestimation of the volume.
  
we already added volume measurements to the TP matches for both the ground truth and predicted piles, summary of that data

```{r}
dbscan_gt_pred_match[["psinf"]] %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::select(field_volume_m3, pred_volume_m3) %>% 
  summary()
```

those don't really look like they match up well...let's explore

```{r}
dbscan_gt_pred_match[["psinf"]] %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::mutate(diff_volume_m3 = field_volume_m3 - pred_volume_m3) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(y = field_volume_m3, x = pred_volume_m3)) +
  ggplot2::geom_abline(lwd = 1.5) +
  # ggplot2::geom_point(ggplot2::aes(color = diff_volume_m3)) +
  ggplot2::geom_point(color = "navy") +
  ggplot2::geom_smooth(method = "lm", se=F, color = "tomato", linetype = "dashed") +
  ggplot2::scale_color_viridis_c(option = "mako", direction = -1, alpha = 0.8) +
  ggplot2::scale_x_continuous(limits = c(0, max( max(dbscan_gt_pred_match[["psinf"]]$pred_volume_m3,na.rm=T), max(dbscan_gt_pred_match[["psinf"]]$field_volume_m3,na.rm=T) ) )) +
  ggplot2::scale_y_continuous(limits = c(0, max( max(dbscan_gt_pred_match[["psinf"]]$pred_volume_m3,na.rm=T), max(dbscan_gt_pred_match[["psinf"]]$field_volume_m3,na.rm=T) ) )) +
  ggplot2::labs(
    y = latex2exp::TeX("allometric field volume $m^3$")
    , x = latex2exp::TeX("predicted volume $m^3$")
    # , color = "image-field\ndiameter diff."
    , subtitle = latex2exp::TeX("bulk volume ($m^3$) comparison")
  ) +
  ggplot2::theme_light()
```

this is exactly what we expected: for true positive matches, there is a clear systematic difference with the plot showing that the volume calculated using the idealized, regular shape assumption (allometric field volume) is consistently larger than the predicted volume derived from the CHM

let's check these using `lm()`

```{r}
lm_temp <- lm(field_volume_m3 ~ pred_volume_m3, data = dbscan_gt_pred_match[["psinf"]] %>% dplyr::filter(match_grp=="true positive"))
summary(lm_temp)
```

These linear model results (intercept = `r scales::comma(lm_temp$coefficients[1], accuracy = 0.01)`, slope = `r scales::comma(lm_temp$coefficients[2], accuracy = 0.01)`) indicate a strong proportional bias that significantly increases with pile size. The high slope (`r scales::comma(lm_temp$coefficients[2], accuracy = 0.01)`) coupled with the negative intercept (`r scales::comma(lm_temp$coefficients[1], accuracy = 0.01)`) indicate that the volume difference is not a simple constant offset (e.g. slope of ~1.0 and intercept of >0 if our hypothesis of consistently higher allometric field volume is true), but rather a scaling issue that is driven by the largest piles. The much larger allometric field volume estimates relative to the CHM-predicted volumes for the largest piles exert a strong influence on the predicted form of the liner model, pulling the slope steeply upward and forcing the intercept below zero as a mathematical artifact. Despite the predicted negative intercept, visual inspection of the data shows that most allometric field volumes are larger than the CHM-predicted volumes, even for smaller piles. The slope value indicates that for every 1 m^3^ increase in predicted volume, the allometric field volume increases by nearly `r scales::comma(lm_temp$coefficients[2], accuracy = 0.01)` m^3^. This data suggests that the geometric assumptions of the allometric model potentially introduce substantial scaling error which may limit its reliability (especially for larger piles) for accurately estimating the volume of real-world piles which have heterogeneous footprints and elevation profiles.

before we compare the volume measurements in aggregate, let's look at the distributions

```{r}
vol_df_temp <- 
  dbscan_gt_pred_match[["psinf"]] %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::select(pile_id,field_volume_m3,pred_volume_m3) %>% 
  tidyr::pivot_longer(cols = -c(pile_id)) %>% 
  dplyr::mutate(
    name = factor(
      name
      , ordered = T
      , levels = c("field_volume_m3","pred_volume_m3")
      , labels = c(
        "allometric field volume"
        , "predicted volume"
      )
    )
  ) 
# plot dist
vol_df_temp %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = value, color = name, fill = name)) +
  ggplot2::geom_density(mapping = ggplot2::aes(y=ggplot2::after_stat(scaled)), alpha = 0.7) +
  harrypotter::scale_color_hp_d(option = "lunalovegood") +
  harrypotter::scale_fill_hp_d(option = "lunalovegood") +
  ggplot2::scale_y_continuous(NULL,breaks=NULL) +
  ggplot2::labs(
    color="",fill="",x=latex2exp::TeX("volume $m^3$")
    , subtitle = latex2exp::TeX("bulk volume ($m^3$) comparison of distributions")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
  )
```

slope plots are neat too

```{r}
vol_df_temp %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = name, y = value, group = pile_id)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  ggplot2::geom_point(mapping = ggplot2::aes(color = name), alpha = 0.7, size = 2.5) +
  harrypotter::scale_color_hp_d(option = "lunalovegood") +
  ggplot2::labs(
    color=""
    , y = latex2exp::TeX("volume $m^3$")
    , x = ""
    , subtitle = latex2exp::TeX("bulk volume ($m^3$) comparison at the pile level")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , axis.title = ggplot2::element_text(size = 10)
    , axis.text = ggplot2::element_text(size = 10)
  )
```

what if we only look at the smaller piles?

```{r}
vol_df_temp %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(x = value < quantile(vol_df_temp$value, probs = 0.90)) %>%
  dplyr::group_by(pile_id) %>% 
  dplyr::filter(
    max(x) == 1
  ) %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = name, y = value, group = pile_id)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  ggplot2::geom_point(mapping = ggplot2::aes(color = name), alpha = 0.7, size = 2.5) +
  harrypotter::scale_color_hp_d(option = "lunalovegood") +
  ggplot2::labs(
    color=""
    , y = latex2exp::TeX("volume $m^3$")
    , x = ""
    , subtitle = latex2exp::TeX("bulk volume ($m^3$) comparison at the pile level for the smaller piles")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , axis.title = ggplot2::element_text(size = 10)
    , axis.text = ggplot2::element_text(size = 10)
  )
```

let's compare aggregated volume measurements for the true positive matches

**Mean Difference (MD):** 
$$\text{MD} = \frac{1}{N} \sum_{i=1}^{N} (\text{Allometric Volume}_i - \text{Predicted Volume}_i)$$

**Percent Mean Difference:** 
$$\%\text{MD} = \frac{\text{MD}}{\text{Mean}(\text{Predicted Volume})} \times 100$$

```{r}
vol_agg_df_temp <- 
  dbscan_gt_pred_match[["psinf"]] %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::ungroup() %>% 
  dplyr::summarise(
    mean_diff = mean(field_volume_m3-pred_volume_m3)
    , sd_diff = sd(field_volume_m3-pred_volume_m3)
    , mean_field_volume_m3 = mean(field_volume_m3,na.rm = T)
    , mean_pred_volume_m3 = mean(pred_volume_m3,na.rm = T)
  ) %>% 
  dplyr::mutate(
    pct_mean_diff = mean_diff/mean_pred_volume_m3
  )
```

what did we get?

```{r}
vol_agg_df_temp %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  dplyr::mutate(
    value = 
      dplyr::case_when(
        stringr::str_starts(name, "pct_") ~ scales::percent(value, accuracy = 0.1)
        , T ~ scales::comma(value, accuracy = 0.1)
      )
  ) %>% 
  kableExtra::kbl(
    caption = "comparison of aggregated allometric field volume and predicted volume"
    , col.names = c("metric", "value")
  ) %>% 
  kableExtra::kable_styling()
```

we'll dig into the MD shortly but before we move on let's focus on the percent mean difference. We calcualted a %MD of `r scales::percent(vol_agg_df_temp$pct_mean_diff, accuracy = 0.1)` which indicates a major systematic difference where the allometric field volume is, on average, `r scales::percent(vol_agg_df_temp$pct_mean_diff, accuracy = 0.1)` larger than our CHM-predicted volume. This large relative difference shows how much the geometric assumptions inflate the volume compared to the irregular volumes measured by our remote sensing-based method.

let's make a Bland-Altman plot to compare the two measurement methods. this plot uses the average of the two measurements (approximate size) on the x-axis and the difference (bias) between the two measurements on the y-axis

```{r}
dbscan_gt_pred_match[["psinf"]] %>% 
  dplyr::filter(match_grp=="true positive") %>%
  dplyr::ungroup() %>% 
  # calc needed metrics
  dplyr::mutate(
    mean_vol = (field_volume_m3+pred_volume_m3)/2
    , diff_vol = (field_volume_m3-pred_volume_m3) # match the order used in vol_agg_df_temp
    , scale_diff = ifelse(diff_vol < 0, -abs(diff_vol) / abs(min(diff_vol)), diff_vol / max(diff_vol))
  ) %>% 
  # ggplot() + geom_point(aes(x=diff_vol,y=0, color=scale_diff)) + scale_color_gradient2(mid = "gray", midpoint = 0, low = "red", high = "blue")
  # plot
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = mean_vol, y = diff_vol)
  ) +
  ggplot2::geom_hline(yintercept = 0, color = "black", lwd = 1.2) +
  # mean difference (bias)
  ggplot2::geom_hline(
    yintercept = vol_agg_df_temp$mean_diff
    , linetype = "dashed", color = "blue", lwd = 1
  ) +
  # upper limit
  ggplot2::geom_hline(
    yintercept = vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff
    , linetype = "dotted", color = "red", lwd = 1
  ) +
  # lower limit
  ggplot2::geom_hline(
    yintercept = vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff
    , linetype = "dotted", color = "red", lwd = 1
  ) +
  # annotations
  ggplot2::annotate(
    "text"
    , x = Inf
    , y = vol_agg_df_temp$mean_diff
    , label = latex2exp::TeX(
      paste0(
        "mean difference (bias): "
        , scales::comma(vol_agg_df_temp$mean_diff, accuracy = 0.01)
        , " $m^3$"
      )
      , output = "character"
    )
    , vjust = -0.5
    , hjust = 1
    , color = "blue"
    , size = 4
    , parse = TRUE
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    , y = vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff
    , label = latex2exp::TeX(
      paste0(
        "+1.96 SD: "
        , scales::comma(vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff, accuracy = 0.01)
        , " $m^3$"
      )
      , output = "character"
    )
    , vjust = -0.5
    , hjust = 1
    , color = "red"
    , size = 4
    , parse = TRUE
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    , y = vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff
    , label = latex2exp::TeX(
      paste0(
        "-1.96 SD: "
        , scales::comma(vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff, accuracy = 0.01)
        , " $m^3$"
      )
      , output = "character"
    )
    , vjust = 1.5
    , hjust = 1
    , color = "red"
    , size = 4
    , parse = TRUE
  ) +
  # points
  ggplot2::geom_point(mapping = ggplot2::aes(color = scale_diff), size = 1.9, alpha = 0.8) +
  ggplot2::scale_color_steps2(mid = "gray", midpoint = 0) +
  ggplot2::labs(
    subtitle = "Bland-Altman plot: allometric field volume vs predicted volume"
    , x = latex2exp::TeX("mean volume ($m^3$)")
    , y = latex2exp::TeX("difference (allometric - predicted volume $m^3$)")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "none")
```

That's a lot of plotting to show that the mean difference is `r scales::comma(vol_agg_df_temp$mean_diff, accuracy = 0.01)` m^3^. Points falling outside the 95% interval on the plot are instances of significant disagreement between the two volume measurements for those specific data points. These outliers indicate that, for a particular pile, the difference between the allometric field volume and the predicted volume is unusually large, suggesting a potential failure in either the CHM segmentation process, the quality of the original field measurements, the geometric shape assumption, or a combination thereof. We should investigate these extreme disagreements further to see what is happening

before we do that, let's use a paired t-test to determine if the mean difference (MD) between the allometric field volume and the predicted volume is statistically significant (i.e. significantly different from zero)

```{r}
# is the mean difference between the two volumes significantly different from zero
ttest_temp <- t.test(
  dbscan_gt_pred_match[["psinf"]] %>% 
    dplyr::filter(match_grp == "true positive") %>% 
    dplyr::pull(field_volume_m3)
  , dbscan_gt_pred_match[["psinf"]] %>% 
    dplyr::filter(match_grp == "true positive") %>% 
    dplyr::pull(pred_volume_m3)
  , paired = TRUE
)
ttest_temp
```

that's neat, the test gave us the same mean difference (MD) of `r scales::comma(vol_agg_df_temp$mean_diff, accuracy = 0.01)` m^3^ that we calculated above. also, the p-value of `r scales::comma(ttest_temp$p.value, accuracy = 0.00001)` is less than 0.05, meaning we should reject the null hypothesis that the true mean difference is zero. this confirms that the systematic difference (or bias) we observed where allometric volume is larger than our predicted volume is statistically significant and not due to random chance.

### Extreme Volume Disagreements

let's investigate the extreme disagreements further to see what is happening 

```{r}
xxx
bad_vol_df_temp <- 
  ground_truth_prediction_match_ans %>% 
  dplyr::filter(match_grp=="true positive") %>%
  dplyr::ungroup() %>% 
  # calc needed metrics
  dplyr::mutate(
    diff_vol = (gt_volume_m3-pred_volume_m3) # match the order used in vol_agg_df_temp
  ) %>% 
  dplyr::filter(
    diff_vol < (vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff)
    | diff_vol > (vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff)
  ) %>% 
  dplyr::left_join(
    slash_piles_polys %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(pile_id, comment) %>% 
      dplyr::rename(pile_type = comment)
    , by = "pile_id"
  )
# what are the differences?
bad_vol_df_temp %>% 
  dplyr::select(
    pile_id, pile_type
    , gt_height_m, pred_height_m, diff_height_m
    , gt_diameter_m, pred_diameter_m, diff_field_diameter_m
    , gt_volume_m3, pred_volume_m3
  ) %>% 
  dplyr::mutate(
    dplyr::across(
      .cols = -c(pile_id,pile_type)
      , .fns = ~ scales::comma(.x,accuracy=0.01)
    )
  ) %>% 
  kableExtra::kbl(
    caption = "Volume measurement outliers: comparison of ground truth and predicted piles"
    # , col.names = c("metric", "value")
  ) %>% 
  kableExtra::kable_styling(font_size = 9.8)
```

All these instances of extreme volume disagreement at the PSINF ponderosa pine training site are specific to mechanical piles and it looks like the field-measured height is much different than the CHM height. This suggests that the optimal `max_ht_m` parameter determined during sensitivity and statistical testing was set too low to properly include the upper extent of these taller machine piles. The parameter was optimized this way because the training data featured a much higher frequency of smaller, shorter hand piles, causing the statistical tuning to be inherently biased toward those dominant smaller objects. This finding highlights the difficulty of creating a single-stage process capable of simultaneously detecting distinct pile forms such as small hand piles and larger machine piles. A workaround for this challenge is to execute the pile detection process in two stages: the first parameterized specifically for the expected size, form, and spectral signature of the hand piles, and a second stage parameterized for the expectations of the larger machine piles and then combine the predictions into a final prediction set.

```{r, fig.height=9.2}
bad_vol_df_temp %>% 
  dplyr::select(
    pile_id
    , gt_height_m, pred_height_m
    , gt_diameter_m, pred_diameter_m
    , gt_volume_m3, pred_volume_m3
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(pile_id)
    , names_to = "metric"
    , values_to = "value"
  ) %>% 
  dplyr::mutate(
    which_data = dplyr::case_when(
        stringr::str_starts(metric,"gt_") ~ "ground 'truth'"
        , stringr::str_starts(metric,"pred_") ~ "prediction"
        , T ~ "error"
      ) %>% 
      ordered()
    , pile_metric = metric %>% 
      stringr::str_remove("(_rmse|_rrmse|_mean|_mape)$") %>% 
      stringr::str_extract("(paraboloid_volume|volume|area|height|diameter)") %>% 
      factor(
        ordered = T
        , levels = c(
          "volume"
          , "paraboloid_volume"
          , "area"
          , "height"
          , "diameter"
        )
        , labels = c(
          "Volume (m3)"
          , "Volume paraboloid"
          , "Area (m2)"
          , "Height (m)"
          , "Diameter (m)"
        )
      )
  ) %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = which_data, y = value, label = scales::comma(value,accuracy=0.1), group = pile_id)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  ggplot2::geom_point(mapping = ggplot2::aes(color = which_data), alpha = 0.8, size = 2.5) +
  ggplot2::scale_color_manual(values = c("blue","brown")) +
  ggplot2::geom_text(
    vjust = -0.25
    , show.legend = FALSE
  ) +
  ggplot2::facet_grid(rows = dplyr::vars(pile_metric), scales = "free_y") +
  ggplot2::scale_y_continuous(labels = scales::comma, expand = ggplot2::expansion(mult = c(0.05,.32))) +
  ggplot2::labs(
    x = "", y = "", color = ""
    , subtitle = "Volume measurement outliers: comparison of measurements"
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , strip.text = ggplot2::element_text(size = 11, color = "black", face = "bold")
  ) 
```

RGB with the predicted piles (brown) and the ground-truth piles (blue)

```{r, fig.height=10.8}
# plot RGB
plts_temp <-
  1:nrow(bad_vol_df_temp) %>% 
  purrr::map(function(x){
    dta <- bad_vol_df_temp %>% dplyr::slice(x)
    gt <- slash_piles_polys %>% dplyr::filter(pile_id==dta$pile_id)
    pr <- final_predicted_slash_piles %>% dplyr::filter(pred_id==dta$pred_id)
    #plt
    ortho_plt_fn(my_ortho_rast=ortho_rast, stand=sf::st_union(gt,pr), buffer=6) +
      ggplot2::geom_sf(data = gt, fill = NA, color = "blue", lwd = 0.6) +
      ggplot2::geom_sf(data = pr, fill = NA, color = "brown", lwd = 0.5)
  })
# combine
patchwork::wrap_plots(
  plts_temp
  , ncol = 2
)
ggplot2::ggsave("../data/PFDP_Data/bad_vol.jpg", height = 8.5, width = 8.5)
```

just looking at the RGB, the pile footprints are in good alignment

let's look at the CHM

```{r, fig.height=10.5, fig.width=8.5}
# cloud2raster_ans$chm_rast %>% 
#   terra::clamp(upper = structural_params_settings$max_ht_m, lower = 0, values = F) %>% 
#   terra::plot()
# bad_vol_df_temp %>% dplyr::glimpse()
# plot RGB + CHM
plts_temp <-
  1:nrow(bad_vol_df_temp) %>% 
  # sample(1) %>% 
  purrr::map(function(x){
    dta <- bad_vol_df_temp %>% dplyr::slice(x)
    gt <- slash_piles_polys %>% dplyr::filter(pile_id==dta$pile_id)
    pr <- final_predicted_slash_piles %>% dplyr::filter(pred_id==dta$pred_id)
    #plt
    ortho_plt_fn(my_ortho_rast=ortho_rast, stand=sf::st_union(gt,pr), buffer=6) +
      ggplot2::geom_tile(
        data = cloud2raster_ans$chm_rast %>% 
          terra::crop(
            sf::st_union(gt,pr) %>% 
              sf::st_transform(terra::crs(cloud2raster_ans$chm_rast)) %>% 
              terra::vect()
          ) %>% 
          terra::mask(
            sf::st_union(gt,pr) %>% 
              sf::st_transform(terra::crs(cloud2raster_ans$chm_rast)) %>% 
              terra::vect()
          ) %>% 
          # slice the chm below our desired height
          # this is what slash_pile_detect_watershed() does
          terra::clamp(upper = structural_params_settings$max_ht_m, lower = 0, values = F) %>%
          terra::as.data.frame(xy=T) %>% 
          dplyr::rename(f=3)
        , mapping = ggplot2::aes(x=x,y=y,fill=f)
        , alpha = 0.5
      ) +
      ggplot2::scale_fill_viridis_c(option = "plasma", na.value = "gray",name = "CHM (m)") +
      ggplot2::geom_sf(data = gt, fill = NA, color = "blue", lwd = 0.6) +
      ggplot2::geom_sf(data = pr, fill = NA, color = "brown", lwd = 0.5) +
      ggplot2::labs(
        subtitle = paste0(
          "GT ht: ", round(dta$gt_height_m,1)
          , " | Pred ht: ", round(dta$pred_height_m,1)
          , "\nGT dia: ", round(dta$gt_diameter_m,1)
          , " | Pred dia: ", round(dta$pred_diameter_m,1)
        )
      ) +
      ggplot2::theme(legend.text = ggplot2::element_text(size = 6),legend.title = ggplot2::element_text(size = 6))
  })
# plts_temp
# combine
patchwork::wrap_plots(
  plts_temp
  , ncol = 2
)
```

The predicted diameters generally align well with the field-measured diameters. However, the field-measured heights for some piles appear unusually high compared to the CHM-derived data.

Our pile detection methodology uses a specific height slice of the CHM based on the `max_ht_m` parameter. This process then keeps only predicted piles that meet the height threshold across the *majority of their area*. The `convexity_pct` parameter functions to filter out predictions where excessive pile top area was removed by the CHM slicing and acts as a safeguard against misclassifying tall objects (like trees) as piles that would have the majority of their top removed by the slicing.

After filtering for expected height, our detection methodology calculates the predicted height as the maximum CHM cell value within the expected height range (i.e. up to the `max_ht_m` setting), which effectively caps any actual height values above this threshold. For our training data, we assumed a maximum height of `r scales::comma(structural_params_settings$max_ht_m,accuracy=0.01)`m. The field data contains heights up to `r scales::comma(max(bad_vol_df_temp$gt_height_m),accuracy=0.01)`m for these volume outliers, which would only be expected of the largest machine piles. In these extreme cases, the significant volume difference observed in the outliers is magnified by both potentially incorrect field measurements and the systematic capping of the predicted height.

To mitigate the risk of vastly underpredicted heights for these tall outliers, one can increase the `convexity_pct` parameter. Making this filtering more strict will remove the outlier height predictions and improve height quantification accuracy for true positive matches, but will likely reduce the detection rate as these piles become false negatives (omissions). Alternatively, one could increase the `max_ht_m` setting to be less restrictive with the height filtering; this would have the effect of estimating a more accurate height value for these tall piles but would potentially introduce new false positive predictions that are actually trees (or yurts?). If spectral data is available to filter these after structural segmentation, then this might be the preferred path to improve height quantification accuracy while maintaining detection accuracy. The choice ultimately depends on whether the user prioritizes detection rate or height quantification accuracy.

There is precedent for smoothing the height of piles from aerial data point cloud data. Trofymow et al. ([2013](https://scholar.google.com/scholar?cluster=3556811352833339764&hl=en&as_sdt=0,6)) calculated pile height as the 95th percentile height of the height-normalized points within the pile polygon. This method was chosen after their preliminary analysis showed it best excluded points from isolated logs extending above the pile while retaining the main pile structure. Our method achieves a functionally similar result for the largest piles by capping the maximum CHM value and smoothing out extreme height measurements.

we can look a the full CHM without capping the height to see how the field-measured values compare to the unfiltered CHM height profile within the pile footprint

```{r, fig.height=10.5, fig.width=8.5}
# cloud2raster_ans$chm_rast %>% 
#   terra::clamp(upper = structural_params_settings$max_ht_m, lower = 0, values = F) %>% 
#   terra::plot()
# bad_vol_df_temp %>% dplyr::glimpse()
# plot RGB + CHM
plts_temp <-
  1:nrow(bad_vol_df_temp) %>% 
  # sample(1) %>% 
  purrr::map(function(x){
    dta <- bad_vol_df_temp %>% dplyr::slice(x)
    gt <- slash_piles_polys %>% dplyr::filter(pile_id==dta$pile_id)
    pr <- final_predicted_slash_piles %>% dplyr::filter(pred_id==dta$pred_id)
    #plt
    ortho_plt_fn(my_ortho_rast=ortho_rast, stand=sf::st_union(gt,pr), buffer=6) +
      ggplot2::geom_tile(
        data = cloud2raster_ans$chm_rast %>% 
          terra::crop(
            sf::st_union(gt,pr) %>% 
              sf::st_transform(terra::crs(cloud2raster_ans$chm_rast)) %>% 
              terra::vect()
          ) %>% 
          terra::mask(
            sf::st_union(gt,pr) %>% 
              sf::st_transform(terra::crs(cloud2raster_ans$chm_rast)) %>% 
              terra::vect()
          ) %>% 
          # DON'T slice the chm below our desired height
          # terra::clamp(upper = structural_params_settings$max_ht_m, lower = 0, values = F) %>%
          terra::as.data.frame(xy=T) %>% 
          dplyr::rename(f=3)
        , mapping = ggplot2::aes(x=x,y=y,fill=f)
        , alpha = 0.5
      ) +
      ggplot2::scale_fill_viridis_c(option = "plasma", na.value = "gray",name = "CHM (m)", breaks = scales::breaks_extended(n=7)) +
      ggplot2::geom_sf(data = gt, fill = NA, color = "blue", lwd = 0.6) +
      ggplot2::geom_sf(data = pr, fill = NA, color = "brown", lwd = 0.5) +
      ggplot2::labs(
        subtitle = paste0(
          "GT ht: ", round(dta$gt_height_m,1)
          , " | Pred ht: ", round(dta$pred_height_m,1)
          , "\nGT dia: ", round(dta$gt_diameter_m,1)
          , " | Pred dia: ", round(dta$pred_diameter_m,1)
        )
      ) +
      ggplot2::theme(legend.text = ggplot2::element_text(size = 6),legend.title = ggplot2::element_text(size = 6))
  })
# plts_temp
# combine
patchwork::wrap_plots(
  plts_temp
  , ncol = 2
)


```

Visual inspection of the unfiltered CHM (note the CHM scale range) within the footprint of these volume outliers confirms that the field-measured values are misaligned with the CHM profile by at least 1m for some of the most extreme outliers, supporting the theory that the significant volume and height differences observed are magnified by potentially incorrect field measurements.

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```


#### Allometric Volume Comparison

The volume comparison [immediately above](#training_detect_final_volcomp) compared the allometric field volume using a geometric shape assumption (paraboloid) with the predicted volume based on the irregular CHM elevation profile. Thus, the volume difference included the impact of the shape assumption. In an attempt to remove the influence of comparing a regular shape (field measured) to an irregular shape (predicted volume), we'll perform a comparison analysis where both volume calculations are based on the same paraboloid shape assumption. This should provide insight into the differences related only to the predicted versus field-measured maximum dimensions of height and diameter.

In the following section, we compare volume predictions against field measured data where both volume calculations utilize the same paraboloid shape assumption. For this analysis, we'll update the predicted volume only using the following volume definitions: 

* **field-measured piles** 
  - **volume** assumes a paraboloid shape, with volume calculated using the field-measured diameter (as the width) and height. we'll refer to this as "Allometric Field Volume" to indicate the field measurement is derived using a shape assumption.
* **predicted piles** 
  - **volume** assumes a paraboloid shape, with volume calculated using the predicted diameter (as the width) and height. we'll refer to this as "Allometric Predicted Volume" to indicate the measurement is derived using a shape assumption.
  
We would generally expect that the allometric field volume will not be uniformly larger or smaller than the predicted allometric volume, as the volume difference will be the net result of two competing influences: the observed `r ifelse(agg_ground_truth_match_ans$diff_field_diameter_m_mean>=0,"overestimation","underestimation")` of diameter (`r scales::comma(agg_ground_truth_match_ans$diff_field_diameter_m_mean,accuracy=0.01)` m mean error) and the observed `r ifelse(agg_ground_truth_match_ans$diff_height_m_mean>=0,"overestimation","underestimation")` of height (`r scales::comma(agg_ground_truth_match_ans$diff_height_m_mean,accuracy=0.01)` m mean error) propagating through the volume calculation formula.

let's first update the predicted volume to calculate the allometric predicted volume where the volume formula for a paraboloid is:
  
$$
V = \frac{1}{8}\pi \cdot width^2 \cdot height
$$
we expect the allometric predicted volume to be larger than the predicted volume calculated based on the irregular elevation profile in the CHM because assuming a geometric shape smooths all the voids and irregularities ( which more accurately reflect reality ;) )

```{r}
ground_truth_prediction_match_ans <- ground_truth_prediction_match_ans %>% 
  dplyr::mutate(
    pred_allom_volume_m3 = (1/8) * pi * (pred_diameter_m^2) * pred_height_m
  )
# summarize it
ground_truth_prediction_match_ans %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::select(gt_volume_m3, pred_volume_m3, pred_allom_volume_m3) %>% 
  summary()
```

nice, the allometric predicted volume is larger than the predicted volume calculated based on the irregular elevation profile in the CHM 

```{r}
ground_truth_prediction_match_ans %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::mutate(diff_volume_m3 = gt_volume_m3 - pred_allom_volume_m3) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(y = gt_volume_m3, x = pred_allom_volume_m3)) +
  ggplot2::geom_abline(lwd = 1.5) +
  ggplot2::geom_point(color = "navy") +
  ggplot2::geom_smooth(method = "lm", se=F, color = "tomato", linetype = "dashed") +
  ggplot2::scale_color_viridis_c(option = "mako", direction = -1, alpha = 0.8) +
  ggplot2::scale_x_continuous(limits = c(0, max( max(ground_truth_prediction_match_ans$pred_allom_volume_m3,na.rm=T), max(ground_truth_prediction_match_ans$gt_volume_m3,na.rm=T) ) )) +
  ggplot2::scale_y_continuous(limits = c(0, max( max(ground_truth_prediction_match_ans$pred_allom_volume_m3,na.rm=T), max(ground_truth_prediction_match_ans$gt_volume_m3,na.rm=T) ) )) +
  ggplot2::labs(
    y = latex2exp::TeX("allometric field volume $m^3$")
    , x = latex2exp::TeX("allometric predicted volume $m^3$")
    # , color = "image-field\ndiameter diff."
    , subtitle = latex2exp::TeX("bulk volume ($m^3$) comparison")
  ) +
  ggplot2::theme_light()
```

the differences between the predictions for the larger field measured piles are pulling the slope of the line upwards but for the smaller piles, the volume differences are clustered around the line of equality

before we compare the volume measurements in aggregate, let's look at the distributions

```{r}
vol_df_temp <- 
  ground_truth_prediction_match_ans %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::select(pile_id,gt_volume_m3,pred_allom_volume_m3) %>% 
  tidyr::pivot_longer(cols = -c(pile_id)) %>% 
  dplyr::mutate(
    name = factor(
      name
      , ordered = T
      , levels = c("gt_volume_m3","pred_allom_volume_m3")
      , labels = c(
        "allometric field volume"
        , "allometric predicted volume"
      )
    )
  ) 
# plot dist
vol_df_temp %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = value, color = name, fill = name)) +
  ggplot2::geom_density(mapping = ggplot2::aes(y=ggplot2::after_stat(scaled)), alpha = 0.7) +
  harrypotter::scale_color_hp_d(option = "lunalovegood") +
  harrypotter::scale_fill_hp_d(option = "lunalovegood") +
  ggplot2::scale_y_continuous(NULL,breaks=NULL) +
  ggplot2::labs(
    color="",fill="",x=latex2exp::TeX("allometric volume $m^3$")
    , subtitle = latex2exp::TeX("Allometric bulk volume ($m^3$) comparison of distributions")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "top"
  )
```

nice, the distributions overlap much more than the volume comparison between the allometric field value and the irregular, CHM-derived prediction

slope plots are neat too

```{r}
vol_df_temp %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = name, y = value, group = pile_id)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  ggplot2::geom_point(mapping = ggplot2::aes(color = name), alpha = 0.7, size = 2.5) +
  harrypotter::scale_color_hp_d(option = "lunalovegood") +
  ggplot2::labs(
    color=""
    , y = latex2exp::TeX("allometric volume $m^3$")
    , x = ""
    , subtitle = latex2exp::TeX("Allometric bulk volume ($m^3$) comparison at the pile level")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , axis.title = ggplot2::element_text(size = 10)
    , axis.text = ggplot2::element_text(size = 10)
  )
```

what if we only look at the smaller piles?

```{r}
vol_df_temp %>% 
  dplyr::filter(
    value < quantile(vol_df_temp$value, probs = 0.938)
  ) %>% 
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = name, y = value, group = pile_id)
  ) +
  ggplot2::geom_line(key_glyph = "point", alpha = 0.7, color = "gray", lwd = 1.1) +
  ggplot2::geom_point(mapping = ggplot2::aes(color = name), alpha = 0.7, size = 2.5) +
  harrypotter::scale_color_hp_d(option = "lunalovegood") +
  ggplot2::labs(
    color=""
    , y = latex2exp::TeX("allometric volume $m^3$")
    , x = ""
    , subtitle = latex2exp::TeX("Allometric bulk volume ($m^3$) comparison at the pile level for the smaller piles")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(
    legend.position = "none"
    , axis.title = ggplot2::element_text(size = 10)
    , axis.text = ggplot2::element_text(size = 10)
  )
```

there is much more variability in this slope plot than the volume comparison between the allometric field value and the irregular, CHM-derived prediction

let's compare aggregated volume measurements for the true positive matches

```{r}
vol_agg_df_temp <- 
  ground_truth_prediction_match_ans %>% 
  dplyr::filter(match_grp=="true positive") %>% 
  dplyr::ungroup() %>% 
  dplyr::summarise(
    mean_diff = mean(gt_volume_m3-pred_allom_volume_m3)
    , sd_diff = sd(gt_volume_m3-pred_allom_volume_m3)
    , mean_gt_volume_m3 = mean(gt_volume_m3,na.rm = T)
    , mean_pred_allom_volume_m3 = mean(pred_allom_volume_m3,na.rm = T)
  ) %>% 
  dplyr::mutate(
    pct_mean_diff = mean_diff/mean_pred_allom_volume_m3
  )
```

what did we get?

```{r}
vol_agg_df_temp %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  dplyr::mutate(
    value = 
      dplyr::case_when(
        stringr::str_starts(name, "pct_") ~ scales::percent(value, accuracy = 0.1)
        , T ~ scales::comma(value, accuracy = 0.1)
      )
  ) %>% 
  kableExtra::kbl(
    caption = "comparison of aggregated allometric field volume and allometric predicted volume"
    , col.names = c("metric", "value")
  ) %>% 
  kableExtra::kable_styling()
```

we'll dig into the MD shortly but before we move on let's focus on the percent mean difference. We calcualted a %MD of `r scales::percent(vol_agg_df_temp$pct_mean_diff, accuracy = 0.1)` which indicates a slight systematic difference where the allometric field volume is, on average, `r scales::percent(vol_agg_df_temp$pct_mean_diff, accuracy = 0.1)` larger than our predicted allometric volume.

let's make a Bland-Altman plot to compare the two measurement methods. this plot uses the average of the two measurements (approximate size) on the x-axis and the difference (bias) between the two measurements on the y-axis

```{r}
ground_truth_prediction_match_ans %>% 
  dplyr::filter(match_grp=="true positive") %>%
  dplyr::ungroup() %>% 
  # calc needed metrics
  dplyr::mutate(
    mean_vol = (gt_volume_m3+pred_allom_volume_m3)/2
    , diff_vol = (gt_volume_m3-pred_allom_volume_m3) # match the order used in vol_agg_df_temp
    , scale_diff = ifelse(diff_vol < 0, -abs(diff_vol) / abs(min(diff_vol)), diff_vol / max(diff_vol))
  ) %>% 
  # ggplot() + geom_point(aes(x=diff_vol,y=0, color=scale_diff)) + scale_color_gradient2(mid = "gray", midpoint = 0, low = "red", high = "blue")
  # plot
  ggplot2::ggplot(
    mapping = ggplot2::aes(x = mean_vol, y = diff_vol)
  ) +
  ggplot2::geom_hline(yintercept = 0, color = "black", lwd = 1.2) +
  # mean difference (bias)
  ggplot2::geom_hline(
    yintercept = vol_agg_df_temp$mean_diff
    , linetype = "dashed", color = "blue", lwd = 1
  ) +
  # upper limit
  ggplot2::geom_hline(
    yintercept = vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff
    , linetype = "dotted", color = "red", lwd = 1
  ) +
  # lower limit
  ggplot2::geom_hline(
    yintercept = vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff
    , linetype = "dotted", color = "red", lwd = 1
  ) +
  # annotations
  ggplot2::annotate(
    "text"
    , x = Inf
    , y = vol_agg_df_temp$mean_diff
    , label = latex2exp::TeX(
      paste0(
        "mean difference (bias): "
        , scales::comma(vol_agg_df_temp$mean_diff, accuracy = 0.01)
        , " $m^3$"
      )
      , output = "character"
    )
    , vjust = -0.5
    , hjust = 1
    , color = "blue"
    , size = 4
    , parse = TRUE
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    , y = vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff
    , label = latex2exp::TeX(
      paste0(
        "+1.96 SD: "
        , scales::comma(vol_agg_df_temp$mean_diff+1.96*vol_agg_df_temp$sd_diff, accuracy = 0.01)
        , " $m^3$"
      )
      , output = "character"
    )
    , vjust = -0.5
    , hjust = 1
    , color = "red"
    , size = 4
    , parse = TRUE
  ) +
  ggplot2::annotate(
    "text"
    , x = Inf
    , y = vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff
    , label = latex2exp::TeX(
      paste0(
        "-1.96 SD: "
        , scales::comma(vol_agg_df_temp$mean_diff-1.96*vol_agg_df_temp$sd_diff, accuracy = 0.01)
        , " $m^3$"
      )
      , output = "character"
    )
    , vjust = 1.5
    , hjust = 1
    , color = "red"
    , size = 4
    , parse = TRUE
  ) +
  # points
  ggplot2::geom_point(mapping = ggplot2::aes(color = scale_diff), size = 1.9, alpha = 0.8) +
  ggplot2::scale_color_steps2(mid = "gray", midpoint = 0) +
  ggplot2::labs(
    subtitle = "Bland-Altman plot: allometric field volume vs allometric predicted volume"
    , x = latex2exp::TeX("mean allometric volume ($m^3$)")
    , y = latex2exp::TeX("difference (allometric field - allometric predicted volume $m^3$)")
  ) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "none")
```

That's a lot of plotting to show that the mean difference is `r scales::comma(vol_agg_df_temp$mean_diff, accuracy = 0.01)` m^3^. This is a significant improvement over our comparison between the allometric field value and the irregular, CHM-derived prediction

let's use a paired t-test to determine if the mean difference (MD) between the allometric field volume and the allometric predicted volume is statistically significant (i.e. significantly different from zero)

```{r}
# is the mean difference between the two volumes significantly different from zero
ttest_temp <- t.test(
  ground_truth_prediction_match_ans %>% 
    dplyr::filter(match_grp == "true positive") %>% 
    dplyr::pull(gt_volume_m3)
  , ground_truth_prediction_match_ans %>% 
    dplyr::filter(match_grp == "true positive") %>% 
    dplyr::pull(pred_allom_volume_m3)
  , paired = TRUE
)
ttest_temp
```

Wow, the p-value of `r scales::comma(ttest_temp$p.value, accuracy = 0.001)` is greater than 0.05 (even 0.10), meaning we *fail* to reject the null hypothesis that the true mean difference is zero. That is, the mean difference between the field-measured volume and the predicted volume is not different than zero when both are forced to use the same paraboloid shape assumption. This contrasts sharply with the previously observed major systematic difference between the allometric field volume and the predicted irregular, CHM-derived volume. The non-significant difference strongly suggests that the major volume discrepancy noted earlier was primarily due to the geometric irregularity of the actual piles, and not to systematic prediction bias in our CHM-derived height and diameter measurements.