# Processing Time{#timer}

A key component of our UAS-based slash pile detection and quantification framework is the cost and time savings compared to traditional field work, especially when used over entire treatment extents to enable census-level monitoring. When we [processed the UAS point cloud](#ptcld_process), we tracked how long it took to generate the CHM raster input required for the proposed framework using the `cloud2trees::cloud2raster()` function. We have compiled the proposed slash pile detection and quantification framework into the `cloud2trees::piles_workflow()` function to enable operational use of the tool. In this section we'll time how long the slash pile methodology takes over the four study sites.


```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# all_stand_boundary %>% dplyr::glimpse()
remove(list = ls()[grep("_temp",ls())])
### delete all other stuff that might conflict with c2t functionality
remove(
  list = setdiff(
    ls()
    , c(
      "chm_rast","rgb_rast","all_stand_boundary"
      , ls()[grep("_out_dir$",ls())]
    )
  )
)
gc()
# chm_rast
# rgb_rast
```

first, we'll rename our `chm_rast` and `rgb_rast` objects in our current R session so that it is clear what we are processing

```{r, results=F, message=F}
chm_rast_list <- chm_rast
rgb_rast_list <- rgb_rast
remove(chm_rast, rgb_rast)
gc()
# organize our out_dir paths where the c2t outputs live as well
  # suffix
  suffix_temp <- "_out_dir"
  # convert to list of objects and remove the suffix from the name in the list
  out_dir_list <-
    all_stand_boundary$site_data_lab %>%
    paste0(suffix_temp) %>%
    base::mget(inherits = T) %>%
    purrr::set_names(~ str_remove(.x, suffix_temp))
  # names(out_dir_list)
  # dplyr::glimpse(out_dir_list)
  # remove
  all_stand_boundary$site_data_lab %>%
    paste0(suffix_temp) %>%
    purrr::map(\(x) remove(list = x,inherits = T))
```

let's see which `cloud2trees` version we have

```{r}
packageVersion("cloud2trees")
```

this version has the `piles_workflow()` function hidden as an internal function so we're going to use `:::` syntax, but future users should be able to use `::` syntax as in `cloud2trees::piles_workflow()`

we'll get the predictions using both the DBSCAN and Watershed segmentation approaches but all we really care about the processing time and not the results themselves since we already generated and evaluated the predictions

```{r, results=F, warning=FALSE}
# read in the cloud2raster() processing times and check if we already ran the piles_workflow()
processed_tracking_data <- 
  out_dir_list %>% 
  purrr::map_dfr(
    \(x)
    dplyr::bind_cols(
      # cloud2raster timer
      readr::read_csv(file = file.path(x,"processed_tracking_data.csv"), progress = F, show_col_types = F)
      # data from las_ctg
      , sf::st_read(dsn = file.path(x,"raw_las_ctg_info.gpkg"), quiet = T) %>% 
        dplyr::ungroup() %>% 
        sf::st_set_geometry("geometry") %>%
        dplyr::summarise(
          geometry = sf::st_union(geometry)
          , number_of_points = sum(Number.of.point.records, na.rm = T)
        ) %>%
        dplyr::mutate(
          las_area_m2 = sf::st_area(geometry) %>% as.numeric()
        ) %>%
        sf::st_drop_geometry()
    )
    , .id = "site_data_lab"
  ) %>% 
  # add calcs
  dplyr::mutate(
    points_per_m2 = number_of_points/las_area_m2
    , timer_cloud2raster_mins_per_ha = timer_cloud2raster_mins/(las_area_m2/10000)
  )
  # dplyr::glimpse()
##################################################
# map over slash_pile_detect for dbscan if haven't done it already
##################################################

out_dir_list %>% 
  purrr::imap(
    function(x, nm){
      # is it already done?
      if(
        any(stringr::str_equal(names(processed_tracking_data), "timer_piles_workflow_mins"))
        && nrow(
            processed_tracking_data %>% 
            dplyr::filter(site_data_lab==nm & !is.na(timer_piles_workflow_mins))
          )==1
      ){
        return()
      }
    }
  )

  # file.names to only run if not already done since it takes a while for all sites
  dir_temp <- "../data"
  if(!dir.exists(dir_temp)){dir.create(dir_temp,showWarnings = F)}
  nmsfx_temp <- "_dbscan_structural_preds.gpkg"
  if(
    !all(file.exists( file.path(dir_temp, paste0(all_stand_boundary$site_data_lab,nmsfx_temp)) ))
  ){
    # map over slash_pile_detect for dbscan
    dbscan_structural_preds <- 
      all_stand_boundary$site_data_lab %>% 
      purrr::set_names() %>% 
      purrr::map(
        \(x)
        slash_pile_detect(
            chm_rast = chm_rast[[x]]
            , seg_method = "dbscan"
            , min_ht_m = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_ht_m)
            , max_ht_m = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(max_ht_m)
            , min_area_m2 = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_area_m2)
            , max_area_m2 = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(max_area_m2)
            , min_convexity_ratio = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_convexity_ratio)
            , min_circularity_ratio = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_circularity_ratio)
            , ofile = file.path(dir_temp, paste0(x,nmsfx_temp))
          )
          # purrr::pluck("segs_sf")
        , .progress = T
      )
    # # write
    #   dbscan_structural_preds %>% 
    #     purrr::imap(
    #       \(x,nm)
    #       sf::st_write(
    #         obj = x
    #         , dsn = file.path(dir_temp, paste0(nm,nmsfx_temp))
    #         , append = F
    #         , quiet = T
    #       )
    #     )
    
    # read files
    dbscan_structural_preds <- dbscan_structural_preds %>% 
      purrr::map(
        \(x)
        sf::st_read(dsn = x, quiet = T)
      )
  }else{
    # read already done
    dbscan_structural_preds <- 
      all_stand_boundary$site_data_lab %>% 
      purrr::set_names() %>% 
      purrr::map(
        \(x)
        sf::st_read(
          dsn = file.path(dir_temp, paste0(x,nmsfx_temp))
          , quiet = T
        )
      )
  }
  # filter all candidate piles for only those that intersect with the study boundary
  dbscan_structural_preds <- 
      all_stand_boundary$site_data_lab %>% 
      purrr::set_names() %>% 
      purrr::map(
        \(x)
        dplyr::inner_join(
          dbscan_structural_preds[[x]]
          , dbscan_structural_preds[[x]] %>% 
            sf::st_intersection(
              stand_boundary[[x]] %>% 
              sf::st_transform(sf::st_crs(dbscan_structural_preds[[x]]))
            ) %>% 
            sf::st_drop_geometry() %>% 
            dplyr::select(pred_id) %>% 
            dplyr::mutate(is_in_stand = T)
          , by = "pred_id"
        )
      )
##################################################
# map over slash_pile_detect for watershed
##################################################
  # file.names to only run if not already done since it takes a while for all sites
  dir_temp <- "../data"
  if(!dir.exists(dir_temp)){dir.create(dir_temp,showWarnings = F)}
  nmsfx_temp <- "_watershed_structural_preds.gpkg"
  if(
    !all(file.exists( file.path(dir_temp, paste0(all_stand_boundary$site_data_lab,nmsfx_temp)) ))
  ){
    # map over slash_pile_detect for watershed
    watershed_structural_preds <- 
      all_stand_boundary$site_data_lab %>% 
      # .[2:4] %>% 
      purrr::set_names() %>% 
      purrr::map(
        \(x)
        slash_pile_detect(
            chm_rast = chm_rast[[x]]
            , seg_method = "watershed"
            , min_ht_m = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_ht_m)
            , max_ht_m = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(max_ht_m)
            , min_area_m2 = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_area_m2)
            , max_area_m2 = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(max_area_m2)
            , min_convexity_ratio = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_convexity_ratio)
            , min_circularity_ratio = all_stand_boundary %>% dplyr::filter(site_data_lab==x) %>% dplyr::pull(min_circularity_ratio)
            , ofile = file.path(dir_temp, paste0(x,nmsfx_temp))
          )
          # purrr::pluck("segs_sf")
        , .progress = T
      )
    # read files
    watershed_structural_preds <- watershed_structural_preds %>% 
      purrr::map(
        \(x)
        sf::st_read(dsn = x, quiet = T)
      )
  }else{
    # read already done
    watershed_structural_preds <- 
      all_stand_boundary$site_data_lab %>% 
      purrr::set_names() %>% 
      purrr::map(
        \(x)
        sf::st_read(
          dsn = file.path(dir_temp, paste0(x,nmsfx_temp))
          , quiet = T
        )
      )
  }
  # filter all candidate piles for only those that intersect with the study boundary
  watershed_structural_preds <- 
      all_stand_boundary$site_data_lab %>% 
      purrr::set_names() %>% 
      purrr::map(
        \(x)
        dplyr::inner_join(
          watershed_structural_preds[[x]]
          , watershed_structural_preds[[x]] %>% 
            sf::st_intersection(
              stand_boundary[[x]] %>% 
              sf::st_transform(sf::st_crs(watershed_structural_preds[[x]]))
            ) %>% 
            sf::st_drop_geometry() %>% 
            dplyr::select(pred_id) %>% 
            dplyr::mutate(is_in_stand = T)
          , by = "pred_id"
        )
      )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

```{r,eval=F,purl=FALSE}
# long stuff
# cloud2trees:::piles_workflow()
```

